### 2024-06-07 draft/wigglyhedra.tex
Updated categorical constraints (lots of todos still).

\documentclass{amsart}

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, multido, xifthen, a4wide, ae, aecompl, blkarray, pifont, mathtools, etoolbox, dsfont}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
\usepackage[bottom]{footmisc}
\usepackage{tikz}
%\usepackage{tkz-graph}
%\usepackage{tikz-qtree}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles, cd}
\usepackage[external]{forest}
%\tikzexternalize
\graphicspath{{figures/}{figures/nodes/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}
\renewcommand{\topfraction}{1} % possibility to have one page of pictures
\renewcommand{\bottomfraction}{1} % possibility to have one page of pictures
\usepackage[noabbrev,capitalise]{cleveref}
\usepackage[export]{adjustbox}
\usepackage{ulem}\normalem

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% theorems
\newtheorem{theorem}{Theorem}%[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem*{theorem*}{Theorem}%[section]

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{question}[theorem]{Question}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{defprop}[theorem]{Definition/Proposition}
\crefname{notation}{Notation}{Notations}
\crefname{problem}{Problem}{Problems}

% math special letters
\newcommand{\R}{\mathbb{R}} % reals
\newcommand{\N}{\mathbb{N}} % naturals
\newcommand{\Z}{\mathbb{Z}} % integers
\newcommand{\C}{\mathbb{C}} % complex
\newcommand{\I}{\mathbb{I}} % set of integers
\newcommand{\HH}{\mathbb{H}} % hyperplane
\newcommand{\K}{\mathbb{K}} % field
\newcommand{\fA}{\mathfrak{A}} % alternating group
\newcommand{\fB}{\mathfrak{S}^\textsc{b}} % signed symmetric group
\newcommand{\cA}{\mathcal{A}} % algebra
\newcommand{\cC}{\mathcal{C}} % collection
\newcommand{\cS}{\mathcal{S}} % ground set
\newcommand{\uR}{\underline{R}} % underline set
\newcommand{\uS}{\underline{S}} % underline set
\newcommand{\uT}{\underline{T}} % underline set
\newcommand{\oS}{\overline{S}} % overline set
\newcommand{\ucS}{\underline{\cS}} % underline ground set
\renewcommand{\c}[1]{\mathcal{#1}} % caligraphic letters
\renewcommand{\b}[1]{{\boldsymbol{#1}}} % bold letters
\newcommand{\bb}[1]{\mathbb{#1}} % bb letters
\newcommand{\f}[1]{\mathfrak{#1}} % frak letters
\newcommand{\h}{\widehat} % hat letters

% math commands
\newcommand{\set}[2]{\left\{ #1 \;\middle|\; #2 \right\}} % set notation
\newcommand{\bigset}[2]{\big\{ #1 \;\big|\; #2 \big\}} % big set notation
\newcommand{\Bigset}[2]{\Big\{ #1 \;\Big|\; #2 \Big\}} % Big set notation
\newcommand{\setangle}[2]{\left\langle #1 \;\middle|\; #2 \right\rangle} % set notation
\newcommand{\ssm}{\smallsetminus} % small set minus
\newcommand{\dotprod}[2]{\left\langle \, #1 \; \middle| \; #2 \, \right\rangle} % dot product
\newcommand{\symdif}{\,\triangle\,} % symmetric difference
\newcommand{\one}{\b{1}} % the all one vector
\newcommand{\eqdef}{\mbox{\,\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}}\ensuremath{=}\,}} % :=
\newcommand{\defeq}{\mbox{~\ensuremath{=}\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}} }} % =:
\newcommand{\simplex}{\b{\triangle}} % simplex
\renewcommand{\implies}{\Rightarrow} % imply sign
\newcommand{\transpose}[1]{{#1}^t} % transpose matrix

% operators
\DeclareMathOperator{\conv}{conv} % convex hull
\DeclareMathOperator{\vect}{vect} % linear span
\DeclareMathOperator{\cone}{cone} % cone hull
\DeclareMathOperator{\inv}{inv} % inversions
\DeclareMathOperator{\ninv}{ninv} % inversions
\DeclareMathOperator{\Ima}{Im} % image
\DeclareMathOperator{\Vol}{Vol} % (mixed) volume
\DeclareMathOperator{\Hom}{Hom} % hom-spaces
\DeclareMathOperator{\Ext}{Ext} % extensions

% others
\newcommand{\ie}{\textit{i.e.}~} % id est
\newcommand{\eg}{\textit{e.g.}~} % exempli gratia
\newcommand{\Eg}{\textit{E.g.}~} % exempli gratia
\newcommand{\apriori}{\textit{a priori}} % a priori
\newcommand{\viceversa}{\textit{vice versa}} % vice versa
\newcommand{\versus}{\textit{vs.}~} % versus
\newcommand{\aka}{\textit{a.k.a.}~} % also known as
\newcommand{\perse}{\textit{per se}} % per se
\newcommand{\ordinal}{\textsuperscript{th}} % th for ordinals
\newcommand{\ordinalst}{\textsuperscript{st}} % st for ordinals
\definecolor{darkblue}{rgb}{0,0,0.7} % darkblue color
\definecolor{green}{RGB}{57,181,74} % darkblue color
\definecolor{violet}{RGB}{147,39,143} % darkblue color
\newcommand{\darkblue}{\color{darkblue}} % darkblue command
\newcommand{\defn}[1]{\textsl{\darkblue #1}} % emphasis of a definition
\newcommand{\para}[1]{\smallskip\noindent\uline{#1.}} % paragraph
\renewcommand{\topfraction}{1} % possibility to have one page of pictures
\renewcommand{\bottomfraction}{1} % possibility to have one page of pictures
%\renewcommand\labelitemi{$\diamond$} % redefine itemize default symbol

% marginal comments
\usepackage{todonotes}
\newcommand{\vincent}[1]{\todo[color=blue!30]{\rm #1 \\ \hfill --- V.}}
\newcommand{\asilata}[2][]{\todo[size=\scriptsize, color=orange!30,#1]{\rm #2 \\ \hfill --- A.}}

% lattices
\newcommand{\meet}{\wedge} % meet
\newcommand{\join}{\vee} % join
\newcommand{\bigMeet}{\bigwedge} % meet
\newcommand{\bigJoin}{\bigvee} % join
\newcommandx{\projDown}[1][1={}]{\smash{\pi_\downarrow^{#1}}} % down projection map
\newcommandx{\projUp}[1][1={}]{\smash{\pi^\uparrow_{#1}}} % up projection map
\newcommand{\con}{\mathrm{con}} % congruence

% geometry
\newcommandx{\Fan}[1][1=D]{\mathcal{F}_{#1}} % fan
\newcommand{\polytope}[1]{\mathds{#1}} % font polytope

% specific wiggly
\newcommand{\wigglyComplex}{\mathrm{WC}} % wiggly complex
\newcommand{\wigglyFlipGraph}{\mathrm{WFG}} % wiggly flip graph
\newcommand{\wigglyIncreasingFlipGraph}{\mathrm{WIFG}} % wiggly increasing flip graph
\newcommand{\wigglyLattice}{\mathrm{WL}} % wiggly lattice
\newcommand{\wigglyFan}{\mathrm{WF}} % wiggly fan
\newcommand{\wigglyhedron}{\polytope{W}} % wigglyhedron
\newcommand{\Asso}{\polytope{A}\mathsf{sso}} % associahedron

% formating the table of contents
\setcounter{tocdepth}{4}
\makeatletter
\def\l@part{\@tocline{1}{8pt}{0pc}{}{}}
\def\l@section{\@tocline{1}{4pt}{0pc}{}{}}
\makeatother
\let\oldtocpart=\tocpart
\renewcommand{\tocpart}[2]{\sc\large\oldtocpart{#1}{#2}}
\let\oldtocsection=\tocsection
\renewcommand{\tocsection}[2]{\bf\oldtocsection{#1}{#2}}
\let\oldtocsubsubsection=\tocsubsubsection
\renewcommand{\tocsubsubsection}[2]{\quad\oldtocsubsubsection{#1}{#2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Wigglyhedra}

\thanks{AB was partially supported by the Australin Research Council DECRA grant DE240100447. 
  VP was partially supported by the Spanish grant PID2022-137283NB-C21 of MCIN/AEI/10.13039/501100011033 / FEDER, UE, by Departament de Recerca i Universitats de la Generalitat de Catalunya (2021 SGR 00697), by the French grant CHARMS (ANR-19-CE40-0017), and by the French--Austrian projects PAGCAP (ANR-21-CE48-0020 \& FWF I 5788).}

\author{Asilata Bapat}
\address{The Australian National University}
\email{asilata.bapat@anu.edu.au}
\urladdr{\url{https://asilata.github.io}}

\author{Vincent Pilaud}
\address{Universitat de Barcelona}
\email{vincent.pilaud@ub.edu}
\urladdr{\url{https://www.ub.edu/comb/vincentpilaud/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{abstract}
\end{abstract}

\maketitle

\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

This problem arose from recent discussions with Asilata Bapat~\cite{BapatPilaud}.
It was originally motivated by her joint work with Anand Deopurkar and Anthony Licata on categorical representation theory and stability conditions, but it can be stated in purely combinatorial terms.

Analogies to cluster algebras and gentle algebras.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section{Wiggly pseudotriangulations and wiggly permutations}
\label{sec:combinatorics}

Fix an integer~$n \ge 1$.
In this section, we define two combinatorial families, the wiggly pseudotriangulations of~$n+2$ points and the wiggly permutations of~$[2n]$, and we prove that they are in bijection.
These two perspectives enable us to endow these combinatorial families with additional structures: the wiggly pseudotriangulations are actually the facets of the wiggly complex, while the wiggly permutations are actually the elements of the wiggly lattice.

%%%%%%%%%%%%

\subsection{Wiggly pseudotriangulations and the wiggly complex}
\label{subsec:wigglyPseudotriangulations}

We start with wiggly pseudodissections of $n+2$ points on a line.
The following definitions are illustrated in \cref{fig:incompatible}.

\begin{definition}
A \defn{wiggly arc} is a quadruple $\alpha \eqdef (i,j,A,B)$ where $0 \le i < j \le n+1$ and the sets~$A$ and~$B$ form a partition of~${]i,j[} \eqdef \{i+1, \dots, j-1\}$.
We represent $\alpha$ by an $x$-monotone curve wiggling around the horizontal axis, starting at point~$i$, ending at point~$j$, and passing above the points of~$A$ and below the points of~$B$.
The wiggly arcs~$\alpha_\mathrm{top} \eqdef (0, n+1, [n], \varnothing)$ and~$\alpha_\mathrm{bot} \eqdef (0, n+1, \varnothing, [n])$ are called \defn{external}, all other wiggly arcs are called \defn{internal}.
\end{definition}

\begin{definition}\label{def:compatible}
Two wiggly arcs~$(i,j,A,B)$ and~$(i',j',A',B')$ are 
\begin{itemize}
\item \defn{non pointed} if~$i = j'$ or~$i' = j$,
\item \defn{crossing} if~$(A \cap B') \cup (\{i,j\} \cap B') \cup (A \cap \{i',j'\}) \ne \varnothing \ne (A' \cap B) \cup (\{i',j'\} \cap B) \cup (A' \cap \{i,j\})$, that is, if the corresponding curves cross,
\item \defn{incompatible} if they are non pointed or crossing, and \defn{compatible} otherwise.
\end{itemize}
%
\begin{figure}[h]
\centerline{\includegraphics[scale=1.3]{incompatible}}
\caption{Some incompatible wiggly arcs: non pointed (left) and crossing (right).}
\label{fig:incompatible}
\end{figure}
\end{definition}

We consider sets of compatible wiggly arcs.
The following definitions are illustrated~in~\cref{fig:pseudodissections}.

\begin{definition}
A \defn{wiggly pseudodissection}~$D$ is a set of pairwise compatible wiggly arcs which contains the exterior wiggly arcs~$\alpha_\mathrm{top}$ and~$\alpha_\mathrm{bot}$. We denote by~$D^\circ \eqdef D \ssm \{\alpha_\mathrm{top}, \alpha_\mathrm{bot}\}$.
%
\begin{figure}[b]
\centerline{\includegraphics[scale=1.5]{pseudodissections}}
\caption{Three wiggly pseudodissections. The left one has three wiggly cells, of respective degrees $3$ (red), $8$ (green), and $4$ (blue). The other two have $7$ pseudotriangles, and are obtained from each other by flipping the wiggly arc separating their yellow and blue pseudotriangles.}
\label{fig:pseudodissections}
\end{figure}
\end{definition}

\vincent{Here, I am using the fact that a set of pairwise non-crossing arcs can be drawn simultaneously non-crossing. I am not sure I want to enter these details.}

\begin{definition}
A wiggly pseudodissection~$D$ decomposes the digon bounded by~$\alpha_\mathrm{top}$ and~$\alpha_\mathrm{bot}$ into cells, that we call \defn{wiggly cells}.
The \defn{degree}~$\delta_c$ of a wiggly cell~$c$ is~$\delta_c \eqdef \beta_c/2+2\kappa_c-1$ where~$\beta_c$ is the number of arcs on the boundary of~$c$ (counted twice if they appear twice along the boundary of~$c$), and~$\kappa_c$ is the number of connected components of the boundary of~$c$.
A point~$p$ on the external boundary of~$c$ is a \defn{corner} (resp.~a \defn{hinge}) of~$c$ if~$c$ covers less (resp.~more) than half of any small disc centered at~$p$.
%The \defn{concave chains} of~$c$ are the maximal sequences of wiggly arcs on the external boundary of~$c$ connected by hinges of~$c$.
\end{definition}

\begin{remark}
\label{rem:degree}
Consider a wiggly cell~$c$ in a wiggly pseudodissection~$D$.
Note that~$\beta_c$ is always even as~$D$ is pointed, so that~$\beta_c/2$ is an integer.
Moreover, we have~$\beta_c \ge 2$ and~$\kappa_c \ge 1$, where at least one of the two inequalities is strict (as otherwise, the two wiggly arcs bounding~$c$ would coincide).
%We conclude that~$\delta_c = \beta_c/2+2\kappa_c-1 \ge 3$ with equality if and only if~$\beta_c = 4$ and~$\kappa_c = 1$.
We thus obtain on the one hand that~$\delta_c = \beta_c/2+2\kappa_c-1 \ge 3$ with equality if and only if~$\beta_c = 4$ and~$\kappa_c = 1$, and on the other hand that~$\beta_c/2+\kappa_c-1 \ge 2$.
\end{remark}

\begin{definition}
A \defn{wiggly pseudotriangle} (resp.~\defn{pseudoquadrangle}) is a wiggly cell of degree~$3$ (resp.~$4$).
\end{definition}

\begin{remark}
\label{rem:descriptionPseudotrianglesPseudoquadrangles}
As illustrated in \cref{fig:pseudotriangles}, a wiggly pseudotriangle~$t$ has no point in its interior and $4$ points on its boundary, $3$ corners and $1$ hinge.
%Hence, it has two corners oriented in the same direction, which we call the \defn{twin corners}.
As illustrated in \cref{fig:pseudoquadrangles}, a wiggly pseudoquadrangle~$q$ can have
\begin{enumerate}
\item either one point in its interior and~$2$ points on its boundary, both corners,
\item or no point in its interior and $6$ points on its boundary, $4$ corners and $2$ hinges. We then distinguish three situations, depending on the position of the two pairs~$p, p'$ of wiggly arcs incident to the two hinges of~$c$, as illustrated in \cref{fig:pseudoquadrangles} (where~$p$ and~$p'$ are colored in red and orange):
\begin{enumerate}[(a)]
\item those where $p$ and $p'$ are not consecutive along~$c$, and do not overlap vertically,
\item those where $p$ and $p'$ are not consecutive along~$c$, but overlap vertically,
\item those where~$p$ and~$p'$ are consecutive along~$c$.
\end{enumerate}
\end{enumerate}
%
\begin{figure}
\centerline{\includegraphics[scale=1.3]{pseudotriangles}}
\caption{Four wiggly pseudotriangles.}
\label{fig:pseudotriangles}
\end{figure}
%
\begin{figure}
\centerline{\includegraphics[scale=1.3]{pseudoquadrangles}}
\vspace{-.3cm}
\caption{Four types of wiggly pseudoquadrangles described in \cref{rem:descriptionPseudotrianglesPseudoquadrangles}.}
\label{fig:pseudoquadrangles}
\end{figure}
\end{remark}

\begin{proposition}
\label{prop:wigglyPseudotriangulations}
Any wiggly pseudodissection~$D$ contains at most~$2n-1$ internal wiggly arcs and at most~$n$ wiggly cells.
Moreover, the following are equivalent:
\begin{enumerate}[(i)]
\item all wiggly cells of~$D$ are wiggly pseudotriangles,
\item $D$ contains~$n$ wiggly cells and is connected,
\item $D$ contains $2n-1$ internal wiggly arcs,
\item $D$ is an inclusion maximal wiggly pseudodissection.
\end{enumerate}
We then say that~$D$ is a \defn{wiggly pseudotriangulation}.
\end{proposition}

\begin{proof}
Consider a wiggly pseudodissection~$D$ with~$\iota$ internal wiggly arcs, $\gamma$ wiggly cells, and~$\kappa$~connected components.
Euler's formula gives~$(n+2)-(\iota+2)+(\gamma+1)=(\kappa+1)$, hence~$\gamma=\iota+\kappa-n$.
Moreover, $\sum_{c \in D} \beta_c/2 = 1+\iota$ since each external (resp.~internal) wiggly arc bounds precisely one (resp.~two) wiggly cell of~$D$, and~${\sum_{c \in D} \big( \kappa_c-1 \big) = \kappa-1}$ since the boundary of each internal connected component of~$D$ is one of the internal connected components of the boundary of precisely one wiggly cell of~$D$.
As~$2 \le \beta_c/2+\kappa_c-1$ for any wiggly cell~$c$ by \cref{rem:degree}, we thus obtain that~$2\gamma \le \sum_{c \in D} (\beta_c/2+\kappa_c-1) = \iota+\kappa = \gamma+n$.
We conclude that~$\gamma \le n$ and~$\iota = \gamma-\kappa+n \le 2n-1$.

%Consider a wiggly pseudodissection~$D$ with~$\iota$ internal wiggly arcs, $\gamma$ wiggly cells, and~$\kappa$~connected components.
%Observe that~$\sum_{c \in D} \beta_c/2 = 1+\iota$ since each external (resp.~internal) wiggly arc bounds precisely one (resp.~two) wiggly cell of~$D$, and that~${\sum_{c \in D} \big( \kappa_c-1 \big) = \kappa-1}$ since the boundary of each internal connected component of~$D$ is one of the internal connected components of the boundary of precisely one wiggly cell of~$D$.
%As~$2 \le \beta_c/2+\kappa_c-1$ for any wiggly cell~$c$ by \cref{rem:degree}, we thus get~$2 \gamma \le \sum_{c \in D} (\beta_c/2+\kappa_c-1) = \iota+\kappa$.
%As~$D$ is a planar graph with~$v = n+2$ vertices, $e = \iota+2$ edges, $f = 1+\gamma$ faces, and $\kappa$ connected components, Euler's formula~$1+\kappa = v-e+f$ yields~$2(1+\kappa) = 2(n+2)-2(\iota+2)+2(1+\gamma) \le 2n-2\iota+2+\iota+\kappa$, so that~$\iota \le 2n-\kappa \le 2n-1$ and~$\gamma \le (\iota+\kappa)/2 \le n$.

To prove the second part of the statement, we show that~(i) $\Rightarrow$ (ii) $\Rightarrow$ (iii) $\Rightarrow$ (iv) $\Rightarrow$ (i).

\para{(i) $\Rightarrow$ (ii)}
For any wiggly cell~$c$, we have~$\delta_c = 3$, so that~$\beta_c = 4$ and~$\kappa_c = 1$ by \cref{rem:degree}, hence~$2 = \beta_c/2+\kappa_c-1$.
Therefore~$\kappa = 1$ ($D$ is connected) and~$2\gamma = \iota+1 = \gamma+n$, so that~$\gamma=n$.

\para{(ii) $\Rightarrow$ (iii)}
As $\gamma=n$ and~$\kappa=1$, we have~$\iota = \gamma+n-\kappa = 2n-1$.

%\para{(i) $\Rightarrow$ (ii)}
%We have~$\delta_c = 3$ for any wiggly cell~$c$.
%Hence~$\beta_c = 4$ and~$\kappa_c = 1$ by \cref{rem:degree}.
%Therefore~$2\gamma = \iota+1$ and~$\kappa = 1$ (so that~$D$ is connected).
%Applying Euler's formula as before, we get~$2 = (n+2)-(\iota+2)+(1+\gamma) = n-\gamma+2$, so that~$\gamma=n$.
%
%\para{(ii) $\Rightarrow$ (iii)}
%As $\gamma=n$ and~$\kappa=1$, we have~$2n = 2\gamma \le \iota+\kappa = \iota+1$, so that~$\iota = 2n-1$.

\para{(iii) $\Rightarrow$ (iv)}
If~$D$ already contains~$2n-1$ internal wiggly arcs, it has maximal cardinality among all wiggly pseudodissections, hence it is certainly inclusion maximal.

\para{(iv) $\Rightarrow$ (i)}
We now consider a wiggly pseudodissection~$D$ containing a wiggly cell~$c$ which is not a wiggly pseudotriangle, and we prove that~$D$ is not inclusion maximal.
%
Assume first that the external boundary of~$c$ is a digon bounded by the wiggly arcs~$\alpha \eqdef (i, j, A, B)$ and~$\alpha' \eqdef (i, j, A', B')$.
Denote by~$m$ the maximum of the points~$(A \symdif A') \cup (B \symdif B')$ located inside this digon (where~$\symdif$ denotes the symmetric difference).
Then the wiggly arc~$\beta \eqdef \big( i, m, {]i,m[} \cap (A \cup A'), {]i,m[} \cap (B \cap B') \big)$ is compatible with~$D$ (any wiggly arc incompatible with~$\beta$ crosses~$\alpha$ or~$\alpha'$) and is not in~$D$ (it connects the boundary of~$c$ with one of its internal connected components).
%Indeed, for any wiggly arc~$\gamma \eqdef (k, \ell, X, Y)$ inside~$c$, we have
%\begin{itemize}
%\item $i < k < \ell \le m$ so that~$\beta$ and~$\gamma$ are pointed,
%\item $Y \subseteq {]i,m[} \cap B \cap B'$ so that~$\beta$ and~$\gamma$ are non-crossing.
%\end{itemize}

Assume now that the external boundary of~$c$ is a not digon.
Consider a hinge~$j$ of~$c$.
Assume by symmetry that the boundary of~$c$ contains two wiggly arcs~$\alpha \eqdef (i, j, A, B)$ and~$\alpha' \eqdef (i', j, A', B')$, such that~$\alpha$ is below~$\alpha'$. % (\ie $k \in A'$ or~$k' \in B$).
Let~$\gamma \eqdef (k, \ell, X, Y)$ denote the uppermost wiggly arc of the boundary of~$c$ such that~$j \in Y$ (it exists since~$j$ is a  hinge of~$c$), and let~$X' \eqdef {]j,\ell[} \cap X$ and~$Y' \eqdef {]j,\ell[} \cap Y$.
Then the two wiggly arcs~$\beta \eqdef \big( i, \ell, A \cup X', B \cup \{j\} \cup Y')$ and~$\beta' \eqdef \big( i', \ell, A'  \cup \{j\} \cup X', B \cup Y')$ are both compatible with~$D$ (any arc incompatible with~$\beta$ is incompatible with~$\alpha$ or~$\gamma$, and similarly for~$\beta'$) and cannot both belong to~$D$ (otherwise, $c$ would be a wiggly pseudotriangle bounded by~$\alpha, \alpha', \gamma', \gamma$).
\end{proof}

\begin{remark}
Note that in the proof of \cref{prop:wigglyPseudotriangulations}, we actually obtained that
\[
\sum_{c \in C} (\delta_c-3) = \sum_{c \in D} \beta_c/2 + 2 \sum_{c \in D} (\kappa_c-1) - 2\gamma = (1+\iota) + 2(\kappa-1) - 2(\iota+\kappa-n) = 2n-1-\iota.
\]
\end{remark}

\begin{definition}
A \defn{wiggly diagonal} of a wiggly cell~$c$ is a wiggly arc inside~$c$ and compatible with~$c$.
\end{definition}

\begin{proposition}
\label{prop:diagonalsPseudoquadrangle}
Any wiggly pseudoquadrangle has exactly two wiggly diagonals.
Moreover, these two wiggly diagonals either cross precisely once, or are non pointed.
\end{proposition}

\begin{proof}
Depending on the four types of pseudoquadrangles described in \cref{rem:descriptionPseudotrianglesPseudoquadrangles}:
\begin{enumerate}
\item the two diagonals are the two wiggly arcs inside~$q$, incident to the point in the interior~of~$q$,
\item[(2a)] the two diagonals connect the two pairs of opposite corners of~$q$,
\item[(2b)] one of two diagonals connects two opposite corners of~$q$ while the other connects the two hinges~of~$q$,
\item[(2c)] one of the diagonals connects two opposite corners of~$q$ while the other connects an hinge of~$q$ to a corner of~$q$.
\end{enumerate}
See \cref{fig:diagonalsPseudoquadrangles} where the two diagonals appear in red and orange.
%
\begin{figure}[h]
\centerline{\includegraphics[scale=1.3]{diagonalsPseudoquadrangles}}
\caption{The two diagonals in each of the four types of wiggly pseudoquadrangles.} % described in \cref{rem:descriptionPseudotrianglesPseudoquadrangles}.}
\label{fig:diagonalsPseudoquadrangles}
\end{figure}
%Let~$q$ be a wiggly pseudoquadrangle.
%We distinguish two situations depending on~$\kappa(q)$:
%\begin{description}
%\item[$\kappa(q) = 1$] Then the boundary of~$c$ is a single wiggly hexagon, so that it has~$4$ concave points. We distinguish again two situations:
%	\begin{description}
%	\item 
%	\end{description}
%\item[$\kappa(q) = 2$] Then~$q$ is a digon bounded by two wiggly arcs~$(i, k, A, B \cup \{j\})$ and~$(i, k, A \cup \{j\}, B)$ for some~$1 < i < j < k \le n$ and~$A \sqcup B = {]i,j[} \cup{]j,k[}$. Hence, the only two wiggly arcs in the interior of~$q$ are~$(i, j, {]i,j[} \cap A, {]i,j[} \cap B)$ and~$(j, k, {]j,k[} \cap A, {]j,k[} \cap B)$, which are both compatible with~$q$ and are not pointed.
%\end{description}
\end{proof}

\vincent{Asilata has an alternative description of the flip...}

We are now ready to consider the wiggly flip graph and the wiggly complex.
The following definitions are illustrated in \cref{fig:wigglyComplex} when~$n = 2$.
%
\begin{figure}
\centerline{\includegraphics[scale=1.1]{wigglyComplex}}
\caption{The wiggly complex~$\wigglyComplex_2$ (left) and the wiggly flip graph~$\wigglyFlipGraph_2$ (right).}
\label{fig:wigglyComplex}
\end{figure}

\begin{definition}
The \defn{wiggly flip graph}~$\wigglyFlipGraph_n$ is the graph with a vertex for each wiggly pseudotriangulation, and with an edge between two wiggly pseudotriangulations~$T,T'$ if there are wiggly arcs~$\alpha \in T$ and~$\alpha' \in T'$ such that~$T \ssm \{\alpha\} = T' \ssm \{\alpha'\}$.
\end{definition}

\begin{proposition}
\label{prop:wigglyFlipGraph}
The wiggly flip graph~$\wigglyFlipGraph_n$ is regular of degree~$2n-1$.
\end{proposition}

\begin{proof}
Consider any internal wiggly arc~$\alpha$ of a wiggly pseudotriangulation~$T$.
Then~$Q \eqdef T \ssm \{\alpha\}$ contains~$n-2$ wiggly pseudotriangles and $1$ pseudoquadrangle~$q$.
By \cref{prop:diagonalsPseudoquadrangle}, $q$ admits precisely two wiggly diagonals, $\alpha$ and another one~$\alpha'$.
As $\alpha'$ is inside~$q$ and compatible with~$c$, it is also compatible with~$Q$.
We obtain that~$T' \eqdef Q \cup \{\alpha'\}$ is the only wiggly pseudotriangulation other than~$T$ containing~$Q$.
\end{proof}

The following definition is illustrated in \cref{fig:incompatible2}.
%
\begin{figure}
\centerline{\includegraphics[scale=1.3]{incompatible2}}
\caption{The arcs~$\alpha$ and~$\alpha'$ of \cref{def:wigglyIncreasingFlipGraph}.}
\label{fig:incompatible2}
\end{figure}

\begin{definition}
\label{def:wigglyIncreasingFlipGraph}
Let~$\alpha \eqdef (i, j, A, B)$ and~$\alpha' \eqdef (i', j', A', B')$ be two wiggly arcs and~$T$ and~$T'$ be two wiggly pseudotriangulations, such that~$T \ssm \{\alpha\} = T' \ssm \{\alpha'\}$.
We say that the flip from~$T$ to~$T'$ is \defn{increasing} if 
\begin{itemize}
\item $\alpha$ starts where~$\alpha'$ ends (that is, $i = j'$), or 
\item $\alpha$ crosses~$\alpha'$ from north-west to south-east (that is, there exists~$k \in [n-1]$ such that $k \in (A \cap B') \cup (\{i\} \cap B') \cup (A \cap \{i'\})$ and~$k+1 \in (A' \cap B) \cup (\{j'\} \cap B) \cup (A' \cap \{j\})$).
\end{itemize}
The \defn{wiggly increasing flip graph}~$\wigglyIncreasingFlipGraph_n$ is the directed graph with a vertex for each wiggly pseudotriangulation and a edge for each increasing~flip.
\end{definition}

\begin{definition}
The \defn{wiggly complex}~$\wigglyComplex_n$ is the simplicial complex of pairwise pointed and non-crossing subsets of internal wiggly arcs.
In other words, it is the clique complex of the compatibility graph on internal wiggly arcs.
\end{definition}

\begin{proposition}
The wiggly complex~$\wigglyComplex_n$ is a flag pure $(2n-1)$-dimensional pseudomanifold without boundary.
\end{proposition}

\begin{proof}
It is a flag simplicial complex because it is a clique complex.
It is pure of dimension $2n-1$ by the implication (iv) $\Rightarrow$ (iii) of \cref{prop:wigglyPseudotriangulations}.
Finally, it is a pseudomanifold without boundary by \cref{prop:wigglyFlipGraph}.
\end{proof}

\begin{remark}
$\wigglyComplex_1$ contains two isolated points.
$\wigglyComplex_2$ is a simplicial $3$-dimensional associahedron (this coincidence fails for~$n > 2$).
Here are the first few $f$-vectors:
\begin{align*}
f(\wigglyComplex_1) & = (1, 2), \\
f(\wigglyComplex_2) & = (1, 9, 21, 14), \\
f(\wigglyComplex_3) & = (1, 24, 154, 396, 440, 176), \\
f(\wigglyComplex_4) & = (1, 55, 729, 4002, 10930, 15684, 11312, 3232), \\
f(\wigglyComplex_5) & = (1, 118, 2868, 28110, 140782, 400374, 673274, 662668, 352728, 78384),
\intertext{and the first few $h$-vectors:}
h(\wigglyComplex_1) & = (1, 1), \\
h(\wigglyComplex_2) & = (1, 6, 6, 1), \\
h(\wigglyComplex_3) & = (1, 19, 68, 68, 19, 1), \\
h(\wigglyComplex_4) & = (1, 48, 420, 1147, 1147, 420, 48, 1), \\
h(\wigglyComplex_5) & = (1, 109, 1960, 11254, 25868, 25868, 11254, 1960, 109, 1).
\end{align*}
\end{remark}

\begin{remark}
Note that the number of vertices of~$\wigglyComplex_n$ is the number~$\sum_{j = 0}^n 2^j (n+1-j)$ of wiggly arcs on~$n+2$ points.
The number of facets of~$\wigglyComplex_n$ is the number of wiggly pseudotriangulations of~$n+2$ points, which is a bit harder to compute.
For this, denote by $f_n(x)$ the polynomial where the coefficient of~$x^i$ is the number of wiggly pseudotriangulations of~$n+2$ points with~$i$ internal wiggly arcs ending at the last point~$n+1$.
For instance,
\begin{align*}
	f_0(x) & = 1, \\
	f_1(x) & = x + 1, \\
	f_2(x) & = 3 x^3 + 5 x^2 + 4 x + 2, \\
	f_3(x) & = 15 x^5 + 35 x^4 + 44 x^3 + 40 x^2 + 28 x + 14, \\
	f_4(x) & = 105 x^7 + 315 x^6 + 520 x^5 + 630 x^4 + 620 x^3 + 514 x^2 + 352 x + 176.
\end{align*}
We use this additional variable~$x$ to obtain a recursive formula for~$f_n(x)$, discussing on the wiggly pseudotriangle~$t_n$ with a hinge at~$n$.
Namely, there are two types of wiggly pseudotriangulations of~$n+2$ points:
\begin{itemize}
\item those with the arc~$(n, n+1, \varnothing, \varnothing)$ are obtained from a wiggly pseudotriangulation of~$n+1$ points by choosing the position of the only corner of~$t_n$ which is not at~$n+1$,
\item those without the arc~$(n, n+1, \varnothing, \varnothing)$ are obtained from a wiggly pseudotriangulation of~$n+1$ points by choosing the positions of the two corners of~$t_n$ which are not at~$n+1$.
\end{itemize}
Hence we obtain for all~$n \ge 1$,
\[
f_{n}(x) = \frac{1}{(1 - x)^2} \big( f_{n-1}(1) + x^2 (2 x^2 - 2 x - 1) f_{n-1}(x) + x^4 (x - 1) f_{n-1}’(x) \big).
\]
This enables to quickly compute the generating series of the number of wiggly pseudotriangulations
\[
\sum_{n \ge 0} f(1) y^n = 1 + 2 y + 14 y^2 + 176 y^3 + 3232 y^4 + 78384 y^5 + 2366248 y^6 + 85534176 y^7 + 3602770400 y^8 + 
% 173300710720 y^9 + 9373542317760 y^10 + 563142033172480 y^11 + 37206559614499840 y^12 + 2681213937595142656 y^13 +
\dots
\]
\end{remark}

\begin{remark}
\label{rem:wigglyComplexAutomorphism}
The vertical and horizontal reflections provide two obvious automorphisms of the wiggly complex~$\wigglyComplex_n$ (and thus of the wiggly flip graph~$\wigglyFlipGraph_n$).
Moreover, the vertical reflection is an anti-automorphism of the wiggly increasing flip graph~$\wigglyIncreasingFlipGraph_n$ (while the horizontal reflection is not).
\end{remark}

Finally, we will need the following observation, illustrated in \cref{fig:incompatible3}.
%
\begin{figure}
\centerline{\includegraphics[scale=1.3]{incompatible3}}
\caption{The arcs~$\beta$ and~$\beta'$ of \cref{prop:uerp}.}
\label{fig:incompatible3}
\end{figure}

\begin{proposition}
\label{prop:uerp}
Let~$\alpha \eqdef (i, j, A, B)$ and~$\alpha' \eqdef (i', j', A', B')$ be two exchangeable wiggly arcs.
Then any wiggly arc compatible with both~$\alpha$ and~$\alpha'$ is also compatible with the wiggly~arcs:
\begin{itemize}
\item $\beta \eqdef (i, j', A \cup A', B \cup B' \cup \{j\})$ and~$\beta' \eqdef (i, j', A \cup A' \cup \{j\}, B \cup B')$ if~$\alpha$ and~$\alpha'$ are non pointed with~$j = i'$,
(In other words, ~$\beta$ and~$\beta'$ are the two wiggly arcs starting at~$\min(i,i')$ and ending at~$\max(j,j')$ which follow $\alpha$ on~$]i,j[$ and~$\alpha'$ on~$]i',j'[$.)
\item $\beta \eqdef \big( i, j', {]i,j'[} \ssm (B \cup B'), {]i,j'[} \cap (B \cup B') \big)$ and~$\beta' \eqdef \big( i', j, {]i',j[} \cap (A \cup A'), {]i',j[} \ssm (A \cup A') \big)$ if~$\alpha$ and~$\alpha'$ are crossing and there is~$k \in [n-1]$ such that~$k \in (A \cap B') \cup (\{i\} \cap B') \cup (A \cap \{i'\})$ and~$k+1 \in (A' \cap B) \cup (\{j'\} \cap B) \cup (A' \cap \{j\})$.
(In other words, $\beta$ is the wiggly arc starting at~$i$ and ending at~$j'$ which follows the lower hull of~$\alpha$ and~$\alpha'$, and~$\beta'$ is the wiggly arc starting at~$i'$ and ending~at~$j$ which follows the upper hull of~$\alpha$ and~$\alpha'$.)
\end{itemize}
%See \cref{fig:incompatible3} for illustrations.
(By exchanging~$\alpha$ and~$\alpha'$, similar statements hold if~$\alpha$ and~$\alpha'$ are non pointed with~$j = i'$, or if~$\alpha$ and~$\alpha'$ are crossing with~$i \in B'$ or~$i' \in A$.)
\end{proposition}

\begin{proof}
Any wiggly arc non pointed with~$\beta$ would be non pointed with~$\alpha$ or~$\alpha'$ (as the endpoints of~$\beta$ are endpoints of~$\alpha$ or~$\alpha'$).
Any wiggly arc crossing~$\beta$ would be either non pointed or crossing~$\alpha$ or~$\alpha'$ (as~$\beta$ follows~$\alpha$ or~$\alpha'$, except at the point~$j = i'$ in the first situation).
\end{proof}

%%%%%%%%%%%%

\subsection{Wiggly permutations and the wiggly lattice}
\label{subsec:wigglyPermutations}

We now consider wiggly permutations of~$[2n]$, defined as follows.

\begin{definition}
A \defn{wiggly permutation} is a permutation of~$[2n]$ which avoids the patterns:
\begin{itemize}
\item $(2j-1) \cdots i \cdots (2j)$ for~$j \in [n]$ and~$i < 2j-1$,
\item $(2j) \cdots k \cdots (2j-1)$ for~$j \in [n]$ and~$k > 2j$.
\end{itemize}
\end{definition}

\begin{example}
The wiggly permutations of~$[2]$ are~$12$ and~$21$.
The wiggly permutations of~$[4]$ are
\[
1234, 2134, 1342, 1243, 3421, 2143, 3412, 1432, 1423, 4321, 4213, 4312, 4132, 4123.
\]
%The wiggly permutations of~$[6]$ are
%\begin{gather*}
%123456, 213456, 134256, 124356, 123564, 123465, 342156, 214356, 213564, 213465, 341256, 143256, \\
%134562, 134265, 142356, 125643, 124365, 135642, 125634, 123654, 123645, 432156, 345621, 342165, \\
%421356, 215643, 214365, 356421, 215634, 213654, 213645, 431256, 341562, 341265, 413256, 143562, \\
%143265, 134652, 134625, 412356, 156423, 142365, 156243, 126543, 126435, 356412, 156342, 136542, \\
%156234, 126534, 126354, 136425, 126345, 435621, 432165, 346521, 346215, 564213, 421365, 562143, \\
%216543, 216435, 563421, 365421, 562134, 216534, 216354, 364215, 216345, 431562, 431265, 345612, \\
%341652, 341625, 413562, 413265, 156432, 143652, 143625, 136452, 564123, 412365, 561423, 165423, \\
%164235, 561243, 165243, 162543, 162435, 563412, 365412, 561342, 165342, 163542, 561234, 165234, \\
%162534, 162354, 364125, 163425, 162345, 564321, 436521, 436215, 364521, 654213, 642135, 652143, \\
%621543, 621435, 653421, 635421, 652134, 621534, 621354, 634215, 621345, 435612, 431652, 431625, \\
%346512, 346152, 346125, 564132, 413652, 413625, 561432, 165432, 164352, 164325, 364152, 163452, \\
%654123, 641235, 651423, 615423, 614235, 651243, 615243, 612543, 612435, 653412, 564312, 635412, \\
%651342, 615342, 613542, 651234, 615234, 612534, 612354, 634125, 613425, 612345, 654321, 643521, \\
%643215, 634521, 436512, 436152, 436125, 364512, 654132, 641352, 641325, 651432, 615432, 614352, \\
%614325, 634152, 613452, 654312, 643125, 643512, 643152, 634512.
%\end{gather*}
\end{example}

\begin{remark}
The numbers of wiggly permutations are given by
\[
\begin{array}{c|ccccccccc}
n & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \dots \\
\hline
|\wigglyLattice_n| & 2 & 14 & 176 & 3232 & 78384 & 2366248 & 85534176 & 3602770400 & \dots
\end{array}
\]
\end{remark}

Recall that the \defn{inversion set} of a permutation~$\sigma$ of~$[2n]$ is the set
\[
\inv(\sigma) \eqdef \set{\big( \sigma(i), \sigma(j) \big)}{1 \le i < j \le 2n \text{ and } \sigma(i) > \sigma(j)}.
\]
We denote by~$\ninv(\sigma)$ the non-inversions of~$\sigma$, that is the complement of~$\inv(\sigma)$.
Note that~$\inv(\sigma)$ and~$\ninv(\sigma)$ are transitive: $\{(k,j), (j,i)\} \subseteq \inv(\sigma)$ implies~$(k,i) \in \inv(\sigma)$ for all~${1 \le i < j < k \le 2n}$, and similarly for~$\ninv$.
The inversion sets of wiggly permutations are characterized as follows.

\begin{lemma}
\label{lem:inversionSetsWigglyPermutations}
A permutation~$\sigma$ of~$[2n]$ is a wiggly permutation if and only if for all~$j \in [n]$,
\begin{itemize}
\item $(2j-1, i) \in \inv(\sigma)$ implies $(2j, i) \in \inv(\sigma)$ for all~$i < 2j-1$, and
\item $(k, 2j-1) \in \inv(\sigma)$ implies $(k, 2j) \in \inv(\sigma)$ for all~$k > 2j$.
\end{itemize}
A similar characterization holds for~$\ninv(\sigma)$ by exchanging~$2j-1$ and~$2j$.
\end{lemma}

\begin{proof}
Immediate as it precisely corresponds to the pattern avoidance description.
\end{proof}

The (left) \defn{weak order} on permutations of~$[2n]$ is defined as the inclusion order of their inversion sets.
Its cover relations are given by the exchanges of two entries at consecutive positions.
It is a (congruence uniform) lattice, and the join and meet satisfy
\[
\inv(\sigma \join \tau) = \big( \inv(\sigma) \cup \inv(\tau) \big)^\textrm{tc}
\qquad\text{and}\qquad
\ninv(\sigma \meet \tau) = \big( \ninv(\sigma) \cup \ninv(\tau) \big)^\textrm{tc}
\]
where~$X^\mathrm{tc}$ denotes the transitive closure of~$X$.

%The following statements are illustrated on \cref{fig:wigglyLattice}\,(right) when~${n = 2}$.
%
\begin{figure}
\centerline{\includegraphics[scale=1.1]{wigglyFlipGraph}}
\caption{The wiggly lattice~$\wigglyLattice_2$ on wiggly pseudotriangulations (left) and on wiggly permutations (right).}
\label{fig:wigglyLattice}
\end{figure}

\begin{proposition}
The wiggly permutations induce a sublattice of the weak order on permutations of~$[2n]$, that we call the \defn{wiggly lattice}~$\wigglyLattice_n$.
\end{proposition}

\begin{proof}
Consider two wiggly permutations~$\rho, \sigma$ of~$[2n]$, and let~$\tau = \rho \join \sigma$.
Assume that there is~$1 \le i < 2j-1 < 2n$ such that~$(2j-1, i) \in \inv(\tau)$.
As $\inv(\tau) = \big( \inv(\rho) \cup \inv(\sigma) \big)^\textrm{tc}$, there exist~$i \le i' < 2j-1$ such that~$(i', i) \in \inv(\tau)$ and~$(2j-1, i') \in \inv(\rho) \cup \inv(\sigma)$.
Since~$\rho$ and~$\sigma$ are wiggly permutations, we obtain by \cref{lem:inversionSetsWigglyPermutations} that~$(2j, i') \in \inv(\rho) \cup \inv(\sigma) \subseteq \inv(\tau)$.
As~$(i', i) \in \inv(\tau)$ and~$(2j, i') \in \inv(\tau)$ and~$\inv(\tau)$ is transitive, we conclude that~$(2j,i) \in \inv(\tau)$.
A similar argument shows that~$(k, 2j-1) \in \inv(\tau)$ implies~$(k, 2j) \in \inv(\tau)$ for all~$1 < 2j < k \le 2n$.
By \cref{lem:inversionSetsWigglyPermutations}, we conclude that~$\tau = \rho \join \sigma$ is a wiggly permutation.
The proof is similar for~$\rho \meet \sigma$, using $\ninv$ instead of~$\inv$.
\end{proof}

\begin{remark}
\label{rem:wigglyLatticeAntiIsomorphism}
The map~$\sigma_1 \dots \sigma_{2n} \mapsto (2n+1-\sigma_1) \dots (2n+1-\sigma_n)$ obviously provides an anti-isomorphism of the wiggly lattice~$\wigglyLattice_n$.
\end{remark}

We now aim at describing the cover relations of the wiggly lattice.
Recall that an \defn{ascent} (resp.~\defn{descent}) in a permutation~$\sigma$ is a position~$j$ such that~$\sigma(j) < \sigma(j+1)$ (resp.~$\sigma(j) > \sigma(j+1)$).

\begin{lemma}\label{lem:wigglyCoverRelation}
Consider an ascent~$j$ of a wiggly permutation~$\sigma$ of~$[2n]$.
Let~$i \eqdef \min \big( j, \sigma^{-1} \big( \sigma(j)+1 \big) \big)$ if~$\sigma(j)$ is odd, and $i \eqdef j$ otherwise.
Let~$k \eqdef \max \big( j+1, \sigma^{-1} \big( \sigma(j+1)+1 \big) \big)$ if~$\sigma(j+1)$ is odd, and $k \eqdef j+1$ otherwise.
Then
\[
\sigma^j \eqdef \sigma(1) \dots \sigma(i-1) \sigma(j+1) \dots \sigma(k) \sigma(i) \dots \sigma(j) \sigma(k+1) \dots \sigma(2n).
\]
is the minimal wiggly permutation such that~$\inv(\sigma) \cup \big\{ \big( \sigma(j+1), \sigma(j) \big) \big\} \subseteq \inv(\sigma^j)$.
\end{lemma}

\begin{proof}
We start with a few observations.
Let~$i < p \le j$ (resp.~$j+1 \le q < k$).
As~$\sigma$ is a wiggly permutation, we have~$\sigma(p) < \sigma(i)$ (resp.~$\sigma(q) > \sigma(k)$).
Hence~$\big( \sigma(i), \sigma(p) \big)$ (resp.~$\big( \sigma(q), \sigma(k) \big)$) belongs to~$\inv(\sigma)$.
Moreover, since~$j$ is an ascent of~$\sigma$, we obtain that~$\sigma(p) < \sigma(q)$ for all~${i \le p \le j < q \le k}$.
Hence, $\inv(\sigma^j) = \inv(\sigma) \cup \bigset{ \big( \sigma(q), \sigma(p) \big) }{i \le p \le j < q \le k}$.

We now prove that~$\sigma^j$ is indeed a wiggly permutation of~$[2n]$.
We consider~$v \in [n]$ and~$u < 2v-1$ such that~$(2v-1, u) \in \inv(\sigma^j)$, and we show that~$(2v, u) \in \inv(\sigma^j)$.
If~$(2v, u) \in \inv(\sigma)$, we are done as~$\inv(\sigma) \subset \inv(\sigma^j)$.
If~$(2v, u) \not\in \inv(\sigma)$, then~$(2v-1, u) \not\in \inv(\sigma)$ by \cref{lem:inversionSetsWigglyPermutations}, and~$i \le \sigma^{-1}(u) \le j < \sigma^{-1}(2v-1) \le k$ since~${\inv(\sigma^j) \ssm \inv(\sigma) = \bigset{ \big( \sigma(q), \sigma(p) \big) }{i \le p \le j < q \le k}}$.
We have~$\sigma^{-1}(2v) \le k$ since~$\sigma$ is a wiggly permutation.
As~$(2v, u) \not\in \inv(\sigma)$ and~$\sigma(p) < \sigma(q)$ for all~$i \le p \le j < q \le k$, we conclude that~$j < \sigma^{-1}(2v) < k$.
Hence, we obtain that ${i \le \sigma^{-1}(u) \le j < \sigma^{-1}(2v) \le k}$, so that~$(2v,u) \in \inv(\sigma^j)$.
Similarly~${(w, 2v-1) \in \inv(\sigma^j)}$ implies~$(w, 2v) \in \inv(\sigma^j)$ for all~$v \in [n]$ and~$w > 2v$.
We conclude that~$\sigma^j$ is indeed a wiggly permutation of~$[2n]$ by \cref{lem:inversionSetsWigglyPermutations}.

Consider now any wiggly transposition~$\tau$ such that~$\inv(\sigma) \cup \big\{ \big( \sigma(j+1), \sigma(j) \big) \big\} \subseteq \inv(\tau)$.
By \cref{lem:inversionSetsWigglyPermutations}, we obtain that~$\inv(\tau)$ also contains % both~$\big( \sigma(j+1), \sigma(i) \big)$ and~$\big( \sigma(k), \sigma(j) \big)$, hence also~
$\big( \sigma(k), \sigma(i) \big)$.
Since~$\inv(\sigma) \subset \inv(\tau)$ and~$\inv(\sigma)$ contains~$\big( \sigma(i), \sigma(p) \big)$ and~$\big( \sigma(q), \sigma(k) \big)$ for all~$i < p \le j$ and~$j+1 \le q < k$, we conclude by transitivity that~$\inv(\tau)$ contains~$\big( \sigma(q), \sigma(p) \big)$ for all~$i \le p \le j < q \le k$.
\end{proof}

\begin{proposition}
Each wiggly permutation~$\sigma$ covers (resp.~is covered by) as many wiggly permutations as its number of descents (resp.~ascents), so that the cover graph of~$\wigglyLattice_n$ is regular of degree~$2n-1$.
\end{proposition}

\begin{proof}
%Consider an ascent~$j$ of~$\sigma$.
%Let~$i \eqdef \min \big( j, \sigma^{-1} \big( \sigma(j)+1 \big) \big)$ if~$\sigma(j)$ is odd, and $i \eqdef j$ otherwise.
%Let~$k \eqdef \max \big( j+1, \sigma^{-1} \big( \sigma(j+1)+1 \big) \big)$ if~$\sigma(j+1)$ is odd, and $k \eqdef j+1$ otherwise.
%Let~$\sigma^j \eqdef \sigma(1) \dots \sigma(i-1) \sigma(j+1) \dots \sigma(k) \sigma(i) \dots \sigma(j) \sigma(k+1) \dots \sigma(2n)$.
%Then~$\sigma^j$ is the minimal wiggly permutation such that~$\inv(\sigma^j) \supseteq \inv(\sigma) \cup \big\{ \big( \sigma(j+1), \sigma(j) \big) \big\}$.
Observe that~$\sigma^j \ne \sigma^{j'}$ for distinct ascents~$j \ne j'$.
As any permutation~$\sigma'$ larger than~$\sigma$ satisfies~${\inv(\sigma') \supseteq \inv(\sigma) \cup \{\sigma(j+1), \sigma(j)\}}$ for some ascent~$j$ of~$\sigma$, we thus obtain that~$j \to \sigma^j$ is a bijection from the ascents of~$\sigma$ to the wiggly permutations covering~$\sigma$.
The proof is symmetrical for the bijection from descents of~$\sigma$ to wiggly permutations covered by~$\sigma$.
\end{proof}

We close this section by the following conjecture (checked computationally up to~$n = 3$).

\begin{conjecture}
\label{conj:Hamiltonian}
The cover graph of the wiggly lattice~$\wigglyLattice_n$ is Hamiltonian.
\end{conjecture}

Note that unfortunately, wiggly permutations do not form a zigzag language in the sense of~\cite{HartungHoangMutzeWilliams} (which would have guarantied the existence of a Gray code).

%%%%%%%%%%%%

\subsection{Bijection}
\label{subsec:bijection}

We now prove that wiggly pseudotriangulations and wiggly permutations are in bijection, and that this bijection induces an isomorphism from the wiggly flip graph to the cover graph of the wiggly lattice.
The following two definitions are illustrated in \cref{fig:wigglyLattice,fig:bijection}.

\begin{figure}
\centerline{\includegraphics[scale=1.7]{bijection}}
\caption{The bijection between wiggly pseudotriangulations and wiggly permutations.}
\label{fig:bijection}
\end{figure}

\begin{definition}
\label{def:bijection1}
Consider a wiggly pseudotriangulation~$T$.
Rotating counterclockwise inside each wiggly pseudotriangle of~$T$, we label by~$2h-1$ (resp.~$2h+1$) the corner immediately preceding (resp.~following) the hinge~$h$ (the remaining corner remains unlabeled).
We define~$\Phi(T)$ as the permutation obtained by reading these labels from bottom to top, meaning that for each internal wiggly arc~$\alpha$ of~$T$, the label incident to~$\alpha$ and below~$\alpha$ appears just before the label incident to~$\alpha$ and above~$\alpha$.
\end{definition}

\begin{definition}
\label{def:bijection2}
For a wiggly permutation~$\sigma$ of~$[2n]$, we define~$\Psi(\sigma) \eqdef \set{\alpha(\sigma, k)}{k \in [2n-1]}$, where for~$k \in [2n-1]$, we have~$\alpha(\sigma, k) \eqdef ( i, j, A, B)$ with
\begin{align*}
i & \eqdef \max \{0\} \cup \set{i \in [n]}{2i-1 \in \sigma([k]) \not\ni 2i}, \\
j & \eqdef \min \{n+1\} \cup \set{j \in [n]}{2j \in \sigma([k]) \not\ni 2j-1}, \\
A & \eqdef \set{\ell \in {]i,j[}}{\{2\ell-1, 2\ell\} \subseteq \sigma([k])}, \\
B & \eqdef \set{\ell \in {]i,j[}}{\{2\ell-1, 2\ell\} \cap \sigma([k]) = \varnothing}.
\end{align*}
\end{definition}

\begin{proposition}
The maps~$\Phi$ of \cref{def:bijection1} and $\Psi$ of \cref{def:bijection2} define inverse bijections between the wiggly pseudotriangulations and the wiggly permutations, which induce a directed graph isomorphism between the wiggly increasing flip graph~$\wigglyIncreasingFlipGraph_n$ and the Hasse diagram of the wiggly lattice~$\wigglyLattice_n$.
\end{proposition}

\begin{proof}
  % We first have to prove that~$\Phi$ and~$\Psi$ are well-defined.
  We break the proof into several steps.

  First, we show that \(\Phi(T)\) is a well-defined wiggly permutation for each wiggly pseudotriangulation \(T\).
  Let \(\sigma = \Phi(T)\), and for any \(j \in [n]\), let \(t_j\) be the pseudotriangle of \(T\) that has hinge \(j\).
  Suppose that \(2j-1\) appears before \(2j\) in \(\sigma\) for some fixed \(j\).
  This means that the pseudotriangle \(t_j\) has its unlabelled corner to the left of its labelled corners.
  In other words, its hinge opens to the right (see, e.g., the triangle labelled by the \(2\)nd hinge in~\cref{fig:bijection}).
  Consider any label \(i\) that we encounter as we traverse from \(2j-1\) to \(2j\).
  This label belongs to some pseudotriangle \(t_h\), such that \(j\) is either a labelled or an unlabelled corner of \(t_h\).
  Moreover, the corner angle of \(t_h\) is sandwiched within the convex angle at \(j\) formed by \(t_j\).
  If \(j\) is an unlabelled corner of \(t_h\), then \(t_h\) also opens to the right at \(h\).
  If \(j\) is a labelled corner of \(t_h\), then \(t_h\) opens to the left.
  In both cases, it is clear that \(h > j\), and hence \(i > 2j-1\).
  In other words, \(\sigma\) avoids the pattern \(2j-1 \cdots i \cdots 2j\) for \(i > 2j-1\).
  The other case (in which \(2j\) appears before \(2j-1\) in \(\sigma\)) is similar.

  Next, we show that \(\Psi(\sigma)\) is a wiggly pseudotriangulation for each wiggly permutation \(\sigma\).
  By~\cref{prop:wigglyPseudotriangulations}, it suffices to check that for each \(k \neq k'\) in \([2n-1]\), the arcs 
  \(\alpha(\sigma,k)\) and \(\alpha(\sigma,k')\) are distinct and compatible.

  To check distinctness, suppose that \(\alpha(\sigma,k) = (i,j,A,B)\) equals \((\sigma,k') = (i',j',A',B')\) for some \(k \leq k'\).
  Note that \(i = i',j = j'\) are both in \(\sigma([k])\), and also \(\sigma([k]) \subseteq \sigma([k'])\).
  For any \(l \in \sigma([k']) \ssm \sigma([k])\), the conditions on wiggly permutations imply that
  \[i < \left\lceil \frac{l}{2} \right\rceil < j.\]
  Furthermore, from the definitions of \(i\) and \(j\), we see that for any \(l \in \sigma([k']) \ssm \sigma([k])\), we have
  \[2 \left\lceil \frac{l}{2} \right\rceil \in \sigma([k']) \text{ if and only if } 2\left\lceil \frac{l}{2} \right\rceil - 1 \in \sigma([k']).\]
  Now it follows from the fact that \(A' = A\) that there is no element \(l \in \sigma([k']) \setminus \sigma([k])\); that is, \(k = k'\).

  We check compatibility.
  Suppose that for some \(k < k'\), we have 
  \[\alpha(\sigma,k) = (i,j,A,B) \text{ and }\alpha(\sigma,k') = (i',j',A',B').\]
  It is clear from the definition that \(i \neq j'\), as the sets over which we take the max and min to define \(i\) and \(j'\) respectively are disjoint.
  Similarly, we have \(j \neq i'\).
  Therefore the two arcs are pointed.
  To check that the arcs are non-crossing, we must show that the sets
  \[(A \cap B') \cup (\{i,j\} \cap B') \cup (A \cap \{i',j'\}) \text{ and } (A' \cap B) \cup (\{i',j'\} \cap B) \cup (A' \cap \{i,j\})\]
  are both empty.
  We check the first; the second is similar.
  Recall that \(\sigma([k]) \subset \sigma([k'])\).
  For any \(l \in B'\), we have \(2l, 2l-1 \notin \sigma([k'])\), and hence \(2l, 2l-1 \notin \sigma([k])\), which means that \(l \notin A\).
  So \(A \cap B' = \emptyset\).
  Since \(2i - 1, 2j \in \sigma([k]) \subset \sigma([k'])\), we see that \(\{i,j\} \cap B' = \emptyset\).
  Since \(2i', 2j'-1 \notin \sigma([k'])\) hence not in \(\sigma([k])\), we see that \(A \cap \{i',j'\} = \emptyset\).

  Finally, we show that the maps \(\Phi\) and \(\Psi\) are inverse bijections.
  Let \(\sigma\) be a wiggly permutation, so that \(\Psi(\sigma)\) is a wiggly pseudotriangulation.
  The following are easy to check for any \(k \in [n]\).
  Let \(m = \lceil \sigma(k)/2 \rceil\).
  \begin{enumerate}
  \item If both \(2m\) and \(2m -1\) appear in \(\sigma([k])\), then \(k\) is the smallest number for which \(\alpha(\sigma,k)\) passes above the point \(m\).
  \item If \(\sigma(k) = 2m\) and \(2m - 1 \notin \sigma([k])\), then \(k\) is the smallest number for which \(\alpha(\sigma,k)\) has its right endpoint at \(m\).
  \item If \(\sigma(k) = 2m - 1\) and \(2m \notin \sigma([k])\), then \(k\) is the smallest number for which \(\alpha(\sigma,k)\) has its left endpoint at \(m\).
  \end{enumerate}
  It follows from the above that in the wiggly pseudotriangulation \(\Psi(\sigma)\), the arc \(\alpha(\sigma,k)\) is precisely the arc lying immediately above the angle \(\sigma(k)\), where we regard \(\alpha(\sigma,2n)\) to be the top irrelevant arc.
  In other words, we have \(\Phi(\Psi(\sigma)) = \sigma\).

  Similarly, let \(T\) be a wiggly pseudotriangulation so that \(\sigma = \Phi(T)\) is a wiggly permutation.
  One can check that for any angle in \(T\), namely any \(l \in [2n]\), the arc immediately above \(l\) is precisely the one constructed as \(\alpha(\sigma,\sigma^{-1}(l))\).
  This shows that \(\Psi(\Phi(T)) = T\).

  Finally, one can check using~\cref{lem:wigglyCoverRelation} that increasing flips in triangulations correspond precisely to edges in the wiggly lattice.
  \asilata{Todo: more details here?}
\item

\end{proof}

\begin{remark}
The map~$\Phi$ sends the wiggly complex automorphism of \cref{rem:wigglyComplexAutomorphism} given by the vertical reflection to the wiggly lattice anti-isomorphism of \cref{rem:wigglyLatticeAntiIsomorphism}.
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Wiggly fan and wigglyhedron}
\label{sec:geometry}

%%%%%%%%%%%%

\subsection{Polyhedral geometry}
\label{subsec:polyhedralGeometry}

We refer to \cite{Ziegler-polytopes} for a reference on polyhedral geometry, and only remind the basic notions needed later in the paper.

A (polyhedral) \defn{cone} is the positive span~$\R_{\ge 0}\b{R}$ of a finite set~$\b{R}$ of vectors of~$\R^d$ or equivalently, the intersection of finitely many closed linear half-spaces of~$\R^d.$ 
The \defn{faces} of a cone are its intersections with its supporting hyperplanes. 
The \defn{rays} (resp.~\defn{facets}) are the faces of dimension~$1$ (resp.~ codimension~$1$).
A cone is \defn{simplicial} if its rays are linearly independent.
A (polyhedral) \defn{fan}~$\Fan$ is a set of cones such that any face of a cone of~$\Fan$ belongs to~$\Fan$, and any two cones of~$\Fan$ intersect along a face of both. 
A fan is \defn{essential} if the intersection of its cones is the origin, \defn{complete} if the union of its cones covers~$\R^d$, and \defn{simplicial} if all its cones are simplicial.

Note that a simplicial fan defines a simplicial complex on its rays (the simplices of the simplicial complex are the subsets of rays which span a cone of the fan).
Conversely, given a simplicial complex~$\Delta$ with ground set~$V$, one can try to realize it geometrically by associating a ray~$\b{r}_v$ of~$\R^d$ to each~$v \in V$, and the cone~$\R_{\ge 0}\b{R}_\triangle$ generated by the set~$\b{R}_\triangle \eqdef \set{\b{r}_v}{v \in \triangle}$ to each~$\triangle \in \Delta$.
To show that the resulting cones indeed form a fan, we will need the following statement, which can be seen as a reformulation of~\cite[Coro.~4.5.20]{DeLoeraRambauSantos}.

\begin{proposition}
\label{prop:characterizationFan}
Consider a closed simplicial $(d-1)$-dimensional pseudomanifold~$\Delta$ with ground set~$V$ and a set of vectors~$(\b{r}_v)_{v \in V}$ of~$\R^d$, and define~$\b{R}_\triangle \eqdef \set{\b{r}_v}{v \in \triangle}$ for any~$\triangle \in \Delta$.
Then the collection of cones~$\set{\R_{\ge 0}\b{R}_\triangle}{\triangle \in \Delta}$ forms a complete simplicial fan of~$\R^d$ if and~only~if
\begin{itemize}
\item there exists a vector~$\b{v}$ of~$\R^d$ contained in only one of the cones~$\R_{\ge 0}\b{R}_\triangle$ for~$\triangle \in \Delta$,
\item for any two adjacent facets~$\triangle, \triangle'$ of~$\Delta$ with~$\triangle \ssm \{v\} = \triangle' \ssm \{v'\}$, we have~$\lambda_v \lambda_{v'} > 0$~where
\[
\lambda_v \, \b{r}_v + \lambda_{v'} \, \b{r}_{v'} + \sum_{w \in \triangle \cap \triangle'} \lambda_w \, \b{r}_w = 0
\]
denotes the unique (up to rescaling) linear dependence on~$\b{R}_{\triangle \cup \triangle'}$.
\end{itemize}
\end{proposition}

A \defn{polytope} is the convex hull of finitely many points of~$\R^d$ or equivalently, a bounded intersection of finitely many closed affine half-spaces of~$\R^d$.
The \defn{faces} of a polytope are its intersections with its supporting hyperplanes.
The \defn{vertices} (resp.~\defn{edges}, resp.~\defn{facets}) are the faces of dimension~$0$ (resp.~dimension~$1$, resp.~codimension~$1$).

The \defn{normal cone} of a face~$\polytope{F}$ of a polytope~$\polytope{P}$ is the cone generated by the normal vectors to the supporting hyperplanes of~$\polytope{P}$ containing~$\polytope{F}$.
Said differently, it is the cone of vectors~$\b{c}$ of~$\R^d$ such that the linear form~$\b{x} \mapsto \dotprod{\b{c}}{\b{x}}$ on~$\polytope{P}$ is maximized by all points of the face~$\polytope{F}$.
The \defn{normal fan} of~$\polytope{P}$ is the set of normal cones of all its faces.

Consider now a complete simplicial fan~$\Fan$ of~$\R^d$ with rays~$(\b{r}_v)_{v \in V}$ and cones~$\R_{\ge 0} \b{R}_\triangle$ for~${\triangle \in \Delta}$, where~$\b{R}_\triangle \eqdef \set{\b{r}_v}{v \in \triangle}$ as in \cref{prop:characterizationFan}.
To realize the fan~$\Fan$, one can try to pick a height vector~$\b{h} \eqdef (h_v)_{v \in V} \in \R^V$ and consider the polytope
\(
\polytope{P}_{\b{h}} \eqdef \set{\b{x} \in \R^d}{\dotprod{\b{r}_v}{\b{x}} \le h_v \text{ for all } v \in V}.
\)
The following classical statement characterizes the height vectors~$\b{h}$ for which the fan~$\Fan$ is the normal fan of this polytope~$\polytope{P}_{\b{h}}$.
We borrow the formulation from~\cite[Lem.~2.1]{ChapotonFominZelevinsky}.

\begin{proposition}
\label{prop:characterizationPolytopalFan}
Let~$\Fan$ be an essential complete simplicial fan in~$\R^n$ with rays~$(\b{r}_v)_{v \in V}$ and cones~$\R_{\ge 0} \b{R}_\triangle$ for~$\triangle \in \Delta$.
Then the following are equivalent for any height vector~$\b{h} \in \R^V$:
\begin{itemize}
\item The fan~$\Fan$ is the normal fan of the polytope~$\polytope{P}_{\b{h}} \eqdef \set{\b{x} \in \R^d}{\dotprod{\b{r}_v}{\b{x}} \le h_v \text{ for all } v \in V}$.
\item For two adjacent facets~$\triangle, \triangle'$ of~$\Delta$ with~$\triangle \ssm \{v\} = \triangle' \ssm \{v'\}$, the height vector~$\b{h}$ satisfies the \defn{wall crossing inequality}
\[
\lambda_v \, h_v + \lambda_{v'} \, h_{v'} + \sum_{w \in \triangle \cap \triangle'} \lambda_w \, h_w > 0
\]
where
\[
\lambda_v \, \b{r}_v + \lambda_{v'} \, \b{r}_{v'} + \sum_{w \in \triangle \cap \triangle'} \lambda_w \, \b{r}_w = 0
\]
denotes the unique linear dependence on~$\b{R}_{\triangle \cup \triangle'}$ such that~$\lambda_v + \lambda_{v'} = 2$.
\end{itemize}
\end{proposition}

We denote by~$(\b{e}_i)_{i \in [d]}$ the standard basis of~$\R^d$.
For~$I \subseteq [d]$, we define~$\one_I \eqdef \sum_{i \in I} \b{e}_i$, and we often shorten~$\one_{[d]}$ by~$\one_d$.
We denote by~$\HH_d$ the hyperplane of~$\R^d$ defined by the equation~$\dotprod{\b{x}}{\one_d} = 0$, and we denote by~$\pi : \R^d \to \HH$ the orthogonal projection on~$\HH_d$, that is~$\pi(\b{x}) \eqdef \b{x} - (\dotprod{\b{x}}{\one_d} / d) \one_d$.

%%%%%%%%%%%%

\subsection{$\b{g}$-vectors and $\b{c}$-vectors}
\label{subsec:gcvectors}

We now define two families of vectors, illustrated in \cref{fig:pseudotriangulationMatrices}.
%
\begin{figure}
\centerline{\raisebox{-1.8cm}{\includegraphics[scale=2]{pseudotriangulationMatrices}} \quad \input{figures/gcmatrices}}
\caption{The $\b{g}$-matrix and $\b{c}$-matrix of a wiggly pseudotriangulation~$T$. The $i$th column of these matrices are the $\b{g}$-vector~$\b{g}(\alpha)$ and $\b{c}$-vector~$\b{c}(\alpha, T)$ of the $i$th wiggly arc~$\alpha$ of~$T$ (ordered from bottom to top). The colors of the columns match the colors of the wiggly arcs.}
\label{fig:pseudotriangulationMatrices}
\end{figure}

\begin{definition}
\label{def:gvectors}
The \defn{$\b{g}$-vector}~$\b{g}(\alpha)$ of a wiggly arc~$\alpha \eqdef (i, j, A, B)$ is defined as the projection~$\pi \big( \hat{\b{g}}(\alpha) \big)$ of the vector~$\hat{\b{g}}(\alpha) \eqdef \one_{\alpha^+} - \one_{\alpha^-}$ where
\begin{align*}
\alpha^+ & \eqdef \big(\{2i-1, 2j\} \ssm \{-1, 2n+2\} \big) \cup \set{2a-1}{a \in A} \cup \set{2a}{a \in A},
\\
\alpha^- & \eqdef \big(\{2i, 2j-1\} \ssm \{0, 2n+1\} \big) \cup \set{2b-1}{b \in B} \cup \set{2b}{b \in B}.
\end{align*}
If we place the coordinate~$2p-1$ on the left and the coordinate~$2p$ on the right of point~$p$, then~$\b{g}(\alpha)$ has a $1$ outside its two endpoints and on both sides of points in~$A$, and a~$-1$ inside its two endpoints and on both sides of points in~$B$.
Note that ${\hat{\b{g}} \big( (0, n+1, [n], \varnothing) \big) = \one_{2n} = - \hat{\b{g}} \big( (0, n+1, \varnothing, [n]) \big)}$, so that~${\b{g} \big( (0, n+1, [n], \varnothing) \big) = \b{0} = \b{g} \big( (0, n+1, \varnothing, [n]) \big)}$.
We define~$\b{g}(D) \eqdef \set{\b{g}(\alpha)}{\alpha \in D^\circ}$ for a set~$D$ of wiggly arcs.
\end{definition}

\begin{definition}
\label{def:cvectors}
The \defn{$\b{c}$-vector} of an interior wiggly arc~$\alpha$ in a wiggly pseudotriangulation~$T$ is the vector~$\b{c}(\alpha, T)$ of~$\HH_{2n}$ such that~$\dotprod{\b{c}(\alpha, T)}{\b{g}(\alpha)} = 1$ and~$\dotprod{\b{c}(\alpha, T)}{\b{g}(\alpha')} = 0$ for all interior wiggly arcs~$\alpha' \ne \alpha$ of~$T$.
In other words, $\b{c}(T) \eqdef \set{\b{c}(\alpha, T)}{\alpha \in T^\circ}$ is the dual basis of~$\b{g}(T)$ in~$\HH_{2n}$.
\end{definition}

\begin{remark}
The $\b{c}$-vector~$\b{c}(\alpha, T)$ can also be described combinatorially.
Namely, label the the corners of~$T$ as in \cref{def:bijection1}, and denote by~$u$ and~$v$ the labels incident to~$\alpha$ and respectively below and above~$\alpha$, and let~$\bar u \eqdef \lceil u/2 \rceil$ and~$\bar v \eqdef \lceil v/2 \rceil$.
Then the $\b{c}$-vector~$\b{c}(\alpha, T)$ has coordinates
\begin{itemize}
\item $\b{c}(\alpha, T)_{2w-1} = - \b{c}(\alpha, T)_{2w} = (-1)^{w \in A} (-1)^{u>v}/4$ for all~$w$ strictly between~$\bar u$ and~$\bar v$,
\item $\b{c}(\alpha, T)_u = 1/2$ if~$\bar u = \bar v$, or if~$\bar v$ and the wiggly arcs incident to~$\bar u$ are on the same side of~$\bar u$, and $\b{c}(\alpha, T)_{2\bar u-1} = \b{c}(\alpha, T)_{2\bar u} = 1/4$ otherwise,
\item $\b{c}(\alpha, T)_v = -1/2$ if~$\bar u = \bar v$, or if~$\bar u$ and the wiggly arcs incident to~$\bar v$ are on the same side of~$\bar v$, and $\b{c}(\alpha, T)_{2\bar v-1} = \b{c}(\alpha, T)_{2\bar v} = -1/4$ otherwise,
\item $\b{c}(\alpha, T)_w = 0$ for all other coordinates.
\end{itemize}
\end{remark}

%\begin{definition}
%\label{def:cvectors}
%Consider a wiggly pseudotriangulation~$T$, and label the corners as in \cref{def:bijection1}.
%Consider an interior wiggly arc~$\alpha$ in a wiggly pseudotriangulation~$T$, denote by~$u$ and~$v$ the labels incident to~$\alpha$ and respectively below and above~$\alpha$, and let~$\bar u \eqdef \lceil u/2 \rceil$ and~$\bar v \eqdef \lceil v/2 \rceil$.
%The \defn{$\b{c}$-vector} of~$\alpha$ in~$T$ is the vector~$\b{c}(\alpha, T)$ with coordinates
%\begin{itemize}
%\item $\b{c}(\alpha, T)_{2w-1} = - \b{c}(\alpha, T)_{2w} = (-1)^{w \in A} (-1)^{u>v}/4$ for all~$w$ strictly between~$\bar u$ and~$\bar v$,
%\item $\b{c}(\alpha, T)_u = 1/2$ if~$\bar u = \bar v$, or if~$\bar v$ and the wiggly arcs incident to~$\bar u$ are on the same side of~$\bar u$, and $\b{c}(\alpha, T)_{2\bar u-1} = \b{c}(\alpha, T)_{2\bar u} = 1/4$ otherwise,
%\item $\b{c}(\alpha, T)_v = -1/2$ if~$\bar u = \bar v$, or if~$\bar u$ and the wiggly arcs incident to~$\bar v$ are on the same side of~$\bar v$, and $\b{c}(\alpha, T)_{2\bar v-1} = \b{c}(\alpha, T)_{2\bar v} = -1/4$ otherwise,
%\item $\b{c}(\alpha, T)_w = 0$ for all other coordinates.
%\end{itemize}
%Set~$\b{c}(T) \eqdef \set{\b{c}(\alpha, T)}{\alpha \in T^\circ}$
%\end{definition}
%
%\begin{proposition}
%\label{prop:dualBases}
%For any wiggly pseudotriangulation~$T$, the sets~$\b{g}(T)$ and~$\b{c}(T)$ are dual bases~of~$\HH_{2n}$.
%\end{proposition}
%
%\begin{proof}
%We claim that~$\dotprod{\hat{\b{g}}(\alpha)}{\b{c}(\alpha', T)} = \delta_{\alpha, \alpha'}$ for~$\alpha, \alpha' \in T$, where~$\delta$ is the Kronecker~$\delta$.
%Note that, with~$\alpha = (0, n+1, [n], \varnothing)$, it implies that~$\dotprod{\one_{2n}}{\b{c}(\alpha', T)} = 0$ for any~${\alpha' \in T}$.
%Hence, we obtain that~$\dotprod{\b{g}(\alpha)}{\b{c}(\alpha', T)} = \dotprod{\pi(\hat{\b{g}}(\alpha))}{\b{c}(\alpha', T)} = \dotprod{\hat{\b{g}}(\alpha)}{\b{c}(\alpha', T)} = \delta_{\alpha, \alpha'}$ for any~$\alpha, \alpha' \in T$.
%We conclude that~$\b{g}(T) \eqdef \set{\b{g}(\alpha)}{\alpha \in T^\circ}$ and~$\b{c}(T) \eqdef \set{\b{c}(\alpha, T)}{\alpha \in T^\circ}$ form dual bases of~$\HH_{2n}$.
%
%The proof of the claim is a boring case analysis.
%Let~$i < j$ be the endpoints of~$\alpha$, and let~$u$ and~$v$ the labels (in the corner labeling of~$T$) incident to~$\alpha'$ and respectively below and above~$\alpha'$, and~$\bar u \eqdef \lceil u/2 \rceil$ and~$\bar v \eqdef \lceil v/2 \rceil$.
%By symmetry, we can assume that~$u < v$ so that~$\bar u \le \bar v$.
%Let us write~$\hat{\b{g}}(\alpha) \eqdef (g_1, \dots, g_{2n})$ and~$\b{c}(\alpha', T) \eqdef (c_1, \dots, c_{2n})$, and define~$gc_k \eqdef g_{2k-1}c_{2k-1} + g_{2k}c_{2k}$.
%Hence, we have~$\dotprod{\hat{\b{g}}(\alpha)}{\b{c}(\alpha', T)} = \sum_{k \in [n]} gc_k = \sum_{k \in \{i, j, \bar u, \bar v\}} gc_k$.
%Indeed, the last equality holds since~$g_{2k-1} = g_{2k}$ for all~$k \in [n] \ssm \{i,j\}$ and~$c_{2k-1} = -c_{2k}$ for all~$k \in [n] \ssm \{\bar u, \bar v\}$.
%
%Assume first that~$\alpha = \alpha'$.
%Hence, we have~$i \le \bar u \le \bar v \le j$.
%We distinguish two cases:
%\begin{enumerate}
%\item If~$i = \bar u = \bar v < j$, then~$u = 2i-1$ and~$v = 2i$, so that~$gc_i = 1 \cdot 1/2 + (-1) \cdot (-1/2) = 1$ and~$gc_j = 0$, thus~$\sum_{k \in \{i, j, \bar u, \bar v\}} gc_k = gc_i + gc_j = 1$.
%\item Otherwise, $i \le \bar u < \bar v \le j$.
%If~$i = \bar u$, we have~$u = 2i-1$ so that~$gc_{\bar u} = 1 \cdot 1/2 + (-1) \cdot 0 = 1/2$.
%If~$i < \bar u$, we have~$gc_i = 0$ (since~$c_{2i-1} = c_{2i} = 0$) and~$gc_{\bar u} = 1/2$ (since~$g_{2\bar u-1} = g_{2\bar u} = 1$ and~${c_{2\bar u-1} + c_{2\bar u} = 1/2}$).
%In both cases, we get that~$\sum_{k \in \{i, \bar u\}} gc_k = 1/2$.
%By symmetric arguments, we conclude~${\sum_{k \in \{i, j, \bar u, \bar v\}} gc_k = \sum_{k \in \{i, \bar u\}} gc_k + \sum_{k \in \{j, \bar v\}} gc_k = 1/2 + 1/2 = 1}$.
%\end{enumerate}
%
%Assume now that~$\alpha \ne \alpha'$.
%\vincent{I am really unhappy with this proof. I can do the case analysis, but it is really unsatisfactory. I am waiting to see if I find another idea...}
%We distinguish ten cases:
%\begin{enumerate}
%\item If~$j < \bar u$, then~$c_{2i-1} = c_{2i} = c_{2j-1} = c_{2j} = 0$ and~$g_{2 \bar u-1} = g_{2 \bar u} = g_{2 \bar v-1} = g_{2 \bar v} = 0$.
%\item If~$j = \bar u < \bar v$, then~$c_{2i-1} \! = \! c_{2i} \! = \! 0$, $g_{2j-1} \! = \! -g_{2j}$ and~$c_{2j-1} \! = \! c_{2j}$,  and~${g_{2 \bar v-1} \! = \! g_{2 \bar v} = 0}$.
%\item If~$i < \bar u < j < \bar v$, then~$c_{2i-1} = c_{2i} = 0$, $g_{2 \bar u-1} = g_{2 \bar u} = 1$ and~$c_{2 \bar u-1} + c_{2 \bar u} = 1/2$, $g_{2j-1} = -g_{2j} = -1$ and~$c_{2j-1} - c_{2j} = 1/2$, and~$g_{2 \bar v-1} = g_{2 \bar v} = 0$.
%\item If~$i < \bar u < j = \bar v$, then~$c_{2i-1} = c_{2i} = 0$, $g_{2 \bar u-1} = g_{2 \bar u} = 1$ and~$c_{2 \bar u-1} + c_{2\bar u} = 1/2$, $g_{2j-1} = -g_{2j} = -1$ and~$c_{2j-1} - c_{2j} = 1/2$.
%\item If~$i < \bar u \le \bar v < j$, then~$c_{2i-1} = c_{2i} = 0$, $g_{2 \bar u-1} = g_{2 \bar u} = g_{2 \bar v-1} = g_{2 \bar v}$ and~$c_{2 \bar u-1} + c_{2\bar u} + c_{2j-1} + c_{2j} = 0$, and~$c_{2j-1} = c_{2j} = 0$.
%\item If~$i = \bar u < j < \bar v$, then \dots
%\item If~$i = \bar u < j = \bar v$, then \dots
%\item If~$i = \bar u < \bar v < j$, then \dots
%\item If~$\bar u < i < j < \bar v$, then \dots
%\item If~$\bar u < i < j = \bar v$, then \dots
%\item If~$\bar u < i < \bar v < j$, then \dots
%\item If~$\bar u < i = \bar v < j$, then \dots
%\item If~$\bar u < \bar v < i < j$, then \dots
%\end{enumerate}
%(Note that~$\bar u = \bar v$ implies that~$\{i,j\} \notin \bar u$.)
%In all cases, we check that~${\sum_{k \in \{i, j, \bar u, \bar v\}} gc_k = 0}$.
%\end{proof}

%%%%%%%%%%%%

\subsection{Wiggly fan}
\label{subsec:wigglyFan}

We now show that the $\b{g}$-vectors of \cref{def:gvectors} support a polyhedral fan realization of the wiggly complex~$\wigglyComplex_n$.

\begin{theorem}
\label{thm:wigglyFan}
The collection of cones~$\R_{\ge 0} \langle \b{g}(A) \rangle$ for all~$A$ in the wiggly complex~$\wigglyComplex_n$ forms a complete simplicial fan of~$\HH_{2n}$, called \defn{wiggly fan}~$\wigglyFan_n$.
\end{theorem}

To prove \cref{thm:wigglyFan}, we just need to check the conditions of \cref{prop:characterizationFan}.
We first find a vector contained in the cone~$\R_{\ge 0} \langle \b{g}(T) \rangle$ for a single wiggly pseudotriangulation~$T$.

\begin{lemma}
\label{lem:-+...-+}
Let~$\b{v} \eqdef \sum_{k \in [n]} (\b{e}_{2k} - \b{e}_{2k-1}) = (-1, 1, \dots, -1, 1)$, let~$\alpha_0 \eqdef (0, 1, \varnothing, \varnothing)$, and for~${i \in [n]}$, let~$\underline{\alpha}_i \eqdef (0, i+1, [i], \varnothing)$ and~$\overline{\alpha}_i \eqdef (0, i+1, \varnothing, [i])$.
Then~$\{\alpha_0\} \cup \bigcup_{i \in [n]} \{\underline{\alpha}_i, \overline{\alpha}_i\}$ is the only wiggly pseudotriangulation~$T$ such that~$\b{v} \in \R_{\ge 0} \langle \b{g}(T) \rangle$.
\end{lemma}

\begin{proof}
Observe first that, for any wiggly arc~$\alpha \eqdef (i, j, A, B)$ and any~$k \in [n]$, we have
%\begin{itemize}
%\item $\hat{\b{g}}(\alpha)_{2k-1} - \hat{\b{g}}(\alpha)_{2k} = 1$ if~$k = i$, $-1$ if~$k = j$, and~$0$ otherwise,
%\item $\hat{\b{g}}(\alpha)_{2k-1} + \hat{\b{g}}(\alpha)_{2k} = 2$ if~$k \in A$, $-2$ if~$k \in B$, and~$0$ otherwise.
%\end{itemize}
\[
\hat{\b{g}}(\alpha)_{2k-1} - \hat{\b{g}}(\alpha)_{2k} = \begin{cases} 1 & \text{if~$k = i$,} \\ -1 & \text{if~$k = j$,} \\ 0 & \text{otherwise}, \end{cases}
\qquad\text{and}\qquad
\hat{\b{g}}(\alpha)_{2k-1} + \hat{\b{g}}(\alpha)_{2k} = \begin{cases} 2 & \text{if~$k \in A$,} \\ -2 & \text{if~$k \in B$,} \\ 0 & \text{otherwise}. \end{cases}
\]

Assume now that~$\b{v} = \sum_\alpha \lambda_\alpha \, \hat{\b{g}}(\alpha)$ where the sum ranges over all wiggly arcs~$\alpha$ for some positive scalars~$\lambda_\alpha \ge 0$ such that~$\lambda_\alpha \cdot \lambda_{\alpha'} = 0$ for all incompatible wiggly arcs~$\alpha$ and~$\alpha'$.
For~$k \in [n]$, let~$\underline{\Lambda}_k$ (resp.~$\overline{\Lambda}_k$) denote the sum of~$\lambda_\alpha$ over the wiggly arcs~$\alpha \eqdef (i, j, A, B)$ with~$k \in A$ (resp.~$k \in B$).
Note that~$\underline{\Lambda}_k = \overline{\Lambda}_k$ since~$0 = \b{v}_{2k-1} + \b{v}_{2k} = \sum_\alpha \lambda_\alpha \big( \hat{\b{g}}(\alpha)_{2k-1} + \hat{\b{g}}(\alpha)_{2k} \big) = 2\underline{\Lambda}_k - 2\overline{\Lambda}_k$.

As~$\b{v}_1 - \b{v}_2 = -2$ and~$\hat{\b{g}}(\alpha)_1 - \hat{\b{g}}(\alpha)_2 \ge 0$ except if~$\alpha = \alpha_0$, we have~$\lambda_{\alpha_0} > 0$, and thus~$\lambda_{\alpha} = 0$ for all wiggly arcs~$\alpha$ starting at~$1$.
%As~$\b{g}(\alpha)_1 - \b{g}(\alpha)_2 = 0$ for all wiggly arcs~$\alpha$ with no endpoint at~$1$, we thus obtain that~$\lambda_{\alpha_0} = 1$.
We then prove by induction on~$i \in [n-1]$ that~$\lambda_{\underline{\alpha}_i} \cdot \lambda_{\overline{\alpha}_i} > 0$ and that~$\lambda_\alpha = 0$ for all wiggly arcs starting at~$i+1$.
Consider~$i \in [n-1]$ and assume that we proved it for~$i-1$.
As~$\smash{\lambda_{\underline{\alpha}_{i-1}} \cdot \lambda_{\overline{\alpha}_{i-1}} > 0}$, we already have~$\lambda_\alpha = 0$ for all wiggly arcs~$\alpha$ ending at~$i+1$, except if~$\alpha \in \{\underline{\alpha}_i, \overline{\alpha}_i\}$.
As~$\b{v}_{2i+1} - \b{v}_{2i+2} = -2$, this implies that~$\smash{\lambda_{\underline{\alpha}_i} + \lambda_{\overline{\alpha}_i} = 1}$.
This already implies that~$\lambda_{\alpha} = 0$ for all wiggly arcs~$\alpha$ starting at~$i+1$.
Moreover, assume for instance that~$\lambda_{\overline{\alpha}_i} > 0$.
Hence, $\lambda_\alpha = 0$ for all wiggly arcs~$\alpha \eqdef (i, j, A, B)$ with~$i \in A$ and~$i+1 \in B$.
We thus obtain that~$\underline{\Lambda}_i - \lambda_{\underline{\alpha}_i} \le \underline{\Lambda}_{i+1}$ and~$\overline{\Lambda}_i - \lambda_{\overline{\alpha}_i} \ge \overline{\Lambda}_{i+1}$.
As~$\underline{\Lambda}_i = \overline{\Lambda}_i$ and~$\underline{\Lambda}_{i+1} = \overline{\Lambda}_{i+1}$, we thus obtain that~$\lambda_{\underline{\alpha}_i} \ge \lambda_{\overline{\alpha}_i} > 0$, hence that~$\lambda_{\underline{\alpha}_i} \cdot \lambda_{\overline{\alpha}_i} > 0$.
We conclude that~$\lambda_\alpha = 0$ for all wiggly arcs~$\alpha$ except if~$\alpha = \alpha_0$ or~$\alpha \in \{\underline{\alpha}_i, \overline{\alpha}_i\}$ for some~$i \in [n]$.

Consider now a wiggly pseudotriangulation~$T$ such that~$\b{v} \in \R_{\ge 0} \langle \b{g}(T) \rangle$.
Write~$\b{v} = \sum_{\alpha \in T^\circ} \lambda_\alpha \, \b{g}(\alpha)$ with~$\lambda_\alpha \ge 0$ for all~$\alpha \in T^\circ$.
Let~$\mu \eqdef \sum_{\alpha \in T^\circ} \lambda_\alpha \dotprod{\hat{\b{g}}(\alpha)}{\one_{2n}}$.
As~$\b{v} = \sum_{\alpha \in T^\circ} \lambda_\alpha \, \hat{\b{g}}(\alpha) - \mu \, \one_{2n}$ and~$\one_{2n} = \hat{\b{g}} \big( (0, n+1, [n], \varnothing) \big) = - \hat{\b{g}} \big( (0, n+1, \varnothing, [n]) \big)$, we thus conclude from the previous paragraph that~$T = \{\alpha_0\} \cup \bigcup_{i \in [n]} \{\underline{\alpha}_i, \overline{\alpha}_i\}$.
\end{proof}

We now describe the unique linear dependence between the $\b{g}$-vectors of two adjacent facets of the wiggly complex~$\wigglyComplex_n$.
The following statement is illustrated in \cref{fig:incompatible3}.

\begin{lemma}
\label{lem:linearDependences}
Let~$T$ and~$T'$ be two adjacent wiggly pseudotriangulations with~$T \ssm \{\alpha\} = T' \ssm \{\alpha'\}$, and let~$\beta$ and~$\beta'$ be the wiggly arcs defined in \cref{prop:uerp}.
Then~$\{\beta, \beta'\} \in T \cap T'$ and
\begin{itemize}
\item if $\alpha$ and~$\alpha'$ are not pointed, then~$\b{g}(\alpha) + \b{g}(\alpha') = \big( \b{g}(\beta) + \b{g}(\beta') \big) / 2$,
\item if~$\alpha$ and~$\alpha'$ are crossing, then~${\b{g}(\alpha) + \b{g}(\alpha') = \b{g}(\beta) + \b{g}(\beta')}$.
\end{itemize}
\end{lemma}

\begin{proof}
In both situations, we will just prove that the same linear dependence holds on the vectors~$\hat{\b{g}}(\alpha), \hat{\b{g}}(\alpha'), \hat{\b{g}}(\beta), \hat{\b{g}}(\beta')$, which will imply the statement as~$\pi$ is a linear map.

If~$\alpha$ and~$\alpha'$ are non pointed, then we have
\[
\alpha^+ \sqcup \alpha'^+ = \beta^+ \sqcup \{2j-1, 2j\} = \beta'^+
\qquad\text{and}\qquad 
\alpha^- \sqcup \alpha'^- = \beta^- = \beta'^- \sqcup \{2j-1, 2j\}.
\]
We thus obtain that
\[
\one_{\alpha^+} + \one_{\alpha'^+} = \one_{\beta^+} + \b{e}_{2j-1} + \b{e}_{2j} = \one_{\beta'^+}
\qquad\text{and}\qquad 
\one_{\alpha^-} + \one_{\alpha'^-} = \one_{\beta^-} = \one_{\beta'^-} + \b{e}_{2j-1} + \b{e}_{2j}.
\]
hence
\begin{align*}
2 \big( \hat{\b{g}}(\alpha) + \hat{\b{g}}(\alpha') \big)
& = 2(\one_{\alpha^+} - \one_{\alpha^-}) + 2(\one_{\alpha'^+} - \one_{\alpha'^-})
= 2(\one_{\alpha^+} + \one_{\alpha'^+}) - 2(\one_{\alpha^-} + \one_{\alpha'^-}) \\
& = (\one_{\beta^+} + \one_{\beta'^+} + \b{e}_{2j-1} + \b{e}_{2j}) - (\one_{\beta^-} + \one_{\beta'^-} + \b{e}_{2j-1} + \b{e}_{2j}) \\
& = (\one_{\beta^+} + \one_{\beta'^+}) - (\one_{\beta^-} + \one_{\beta'^-})
= (\one_{\beta^+} - \one_{\beta^-}) + (\one_{\beta^-} + \one_{\beta'^-})
= \hat{\b{g}}(\beta) + \hat{\b{g}}(\beta').
\end{align*}

If~$\alpha$ and~$\alpha'$ cross, then we have
\begin{gather*}
\alpha^+ \cup \alpha'^+ = \beta^+ \cup \beta'^+ \qquad\text{and}\qquad \alpha^+ \cap \alpha'^+ = \beta^+ \cap \beta'^+, \\
\alpha^- \cup \alpha'^- = \beta^- \cup \beta'^- \qquad\text{and}\qquad \alpha^- \cap \alpha'^- = \beta^- \cap \beta'^-.
\end{gather*}
We thus obtain that
\[
\one_{\alpha^+} + \one_{\alpha'^+} = \one_{\alpha^+ \cup \alpha'^+} + \one_{\alpha^+ \cap \alpha'^+} = \one_{\beta^+ \cup \beta'^+} + \one_{\beta^+ \cap \beta'^+} = \one_{\beta^+} + \one_{\beta'^+}
\]
and similarly~$\one_{\alpha^-} + \one_{\alpha'^-} = \one_{\beta^-} + \one_{\beta'^-}$, so that~$\hat{\b{g}}(\alpha) + \hat{\b{g}}(\alpha') =\hat{\b{g}}(\beta) + \hat{\b{g}}(\beta')$.
\end{proof}

\begin{proof}[Proof of \cref{thm:wigglyFan}]
Directly follows from \cref{prop:characterizationFan,lem:-+...-+,lem:linearDependences}.
\end{proof}

\begin{remark}
The wiggly fan coarsens the $\b{c}$-vector fan.
\vincent{todo}
\end{remark}

%%%%%%%%%%%%

\subsection{Wigglyhedron}
\label{subsec:wigglyhedron}

We now construct the wigglyhedron illustrated in \cref{fig:wigglyhedron} when~${n = 2}$.
%
\begin{figure}[t]
\centerline{\input{wigglyhedron}}
\caption{The wigglyhedron~$\wigglyhedron_2$.}
\label{fig:wigglyhedron}
\end{figure}

\begin{definition}
\label{def:incompatibilityDegree}
The \defn{incompatibility degree}~$\delta(\alpha, \alpha')$ of two wiggly arcs~$\alpha \eqdef (i, j, A, B)$ and~$\alpha' \eqdef (i', j', A', B')$ is given by
\begin{itemize}
\item $0$ if~$\alpha$ and~$\alpha'$ are pointed and non-crossing,
\item $1$ is~$\alpha$ and~$\alpha'$ are not pointed (\ie $i = j'$ or~$i' = j$),
\item the number of crossings of~$\alpha$ and~$\alpha'$ if they are crossing (\ie the number of~$p \in \{0, \dots, n\}$ with $p \in (A \cap B') \cup (\{i,j\} \cap B') \cup (A \cap \{i',j'\})$ and~$p+1 \in (A' \cap B) \cup (\{i',j'\} \cap B) \cup (A' \cap \{i,j\})$, or the opposite).
\end{itemize}
The \defn{incompatibility number} of a wiggly arc~$\alpha$ is~$\kappa(\alpha) \eqdef \sum_{\alpha'} \delta(\alpha, \alpha')$.
\end{definition}

\begin{remark}
Note that~$\delta$ is indeed an incompatibility degree in the sense that
\begin{itemize}
\item $\alpha$ and~$\alpha'$ are compatible if and only if~$\delta(\alpha, \alpha') = 0$,
\item $\alpha$ and~$\alpha'$ are exchangeable if and only if~$\delta(\alpha, \alpha') = 1$.
\end{itemize}
Observe also that~$\delta$ (and thus~$\kappa$) is preserved by the automorphisms of \cref{rem:wigglyComplexAutomorphism}.
\end{remark}

\vincent{The incompatibility degree is the number of ``close'' minors of the form~$\begin{bmatrix} 1 & -1 \\ -1 & 1 \end{bmatrix}$ or~$\begin{bmatrix} -1 & 1 \\ 1 & -1 \end{bmatrix}$ in the matrix obtained from the two $\b{g}$-vectors of the wiggly arcs.}

\begin{example}
Observe that
$\kappa \big( (0, j, {[1,j[}, \varnothing) \big) 
= \kappa \big( (0, j, \varnothing, {[1,j[}) \big) 
%= \kappa \big( (n+1-j, n+1, {]n+1-j, n]}, \varnothing) \big) 
%= \kappa \big( (n+1-j, n+1, \varnothing, {]n+1-j, n]}) \big) 
= (2^j-1)(2^{n+1-j}-1)$.
\end{example}

\begin{lemma}
\label{lem:wallCrossingInequalities}
Let~$T$ and~$T'$ be two adjacent wiggly pseudotriangulations with~$T \ssm \{\alpha\} = T' \ssm \{\alpha'\}$, and let~$\beta$ and~$\beta'$ be the wiggly arcs defined in \cref{prop:uerp}.
Then
\begin{itemize}
\item if $\alpha$ and~$\alpha'$ are not pointed, then~$\kappa(\alpha) + \kappa(\alpha') > \big( \kappa(\beta) + \kappa(\beta') \big) / 2$,
\item if~$\alpha$ and~$\alpha'$ are crossing, then~${\kappa(\alpha) + \kappa(\alpha') > \kappa(\beta) + \kappa(\beta')}$.
\end{itemize}
\end{lemma}

\begin{proof}
Any crossing between a wiggly arc~$\gamma$ and~$\beta$ or~$\beta'$ translates to a crossing with~$\alpha$ or~$\alpha'$.
When~$\alpha$ and~$\alpha'$ are not pointed, a crossing with~$\alpha$ or~$\alpha'$ can correspond to a crossing with~$\beta$ and a crossing with~$\beta'$, so that~$\delta(\alpha, \gamma) + \delta(\alpha', \gamma) \ge \big( \delta(\beta, \gamma) + \delta(\beta', \gamma) \big) / 2$.
When~$\alpha$ and~$\alpha'$ are crossing, then there is a correspondence between the crossings with~$\alpha$ or~$\alpha'$ and the crossings with~$\beta$ or~$\beta'$, so that~$\delta(\alpha, \gamma) + \delta(\alpha', \gamma) \ge \delta(\beta, \gamma) + \delta(\beta', \gamma)$
The result follows by summation, and since~$\delta(\alpha, \alpha') = 1$.
\end{proof}

\begin{theorem}
\label{thm:wigglyhedron}
The wiggly fan~$\wigglyFan_n$ is the normal fan of a simplicial $(2n-1)$-dimensional polytope, called the \defn{wigglyhedron}~$\wigglyhedron_n$, and defined equivalently~as
\begin{itemize}
\item the intersection of the halfspaces~$\set{\b{x} \in \HH}{\dotprod{\b{g}(\alpha)}{\b{x}} \le \kappa(\alpha)}$ for all internal wiggly arcs~$\alpha$,
\item the convex hull of the points~$\b{p}(T) \eqdef \sum\limits_{\alpha \in T} \kappa(\alpha) \, \b{c}(\alpha, T)$ for all wiggly pseudotriangulations~$T$.
\end{itemize}
\end{theorem}

\begin{proof}
By \cref{prop:characterizationPolytopalFan,lem:linearDependences,lem:wallCrossingInequalities}, the wiggly fan is indeed the normal fan of the polytope~$\wigglyhedron_n$ defined by the inequalities~$\dotprod{\b{g}(\alpha)}{\b{x}} \le \kappa(\alpha)$ for all internal wiggly arcs~$\alpha$.
For any wiggly pseudotriangulation~$T$ and any wiggly arc~$\alpha$ of~$T$, we have~$\dotprod{\b{g}(\alpha)}{\b{p}(T)} = \sum_{\alpha' \in T} \kappa(\beta) \dotprod{\b{g}(\alpha)}{\b{c}(\alpha', T)} = \kappa(\alpha)$ since~$\b{g}(T)$ and~$\b{c}(T)$ are dual bases by \cref{def:cvectors}.
Hence, $\b{p}(T)$ is indeed the point of~$\wigglyhedron_n$ corresponding to~$T$, which is located at the intersection of the hyperplanes~$\dotprod{\b{g}(\alpha)}{\b{x}} = \kappa(\alpha)$ for all internal wiggly arcs~$\alpha$ in~$T$.
\end{proof}

In fact, the wigglyhedron even recovers the wiggly lattice of \cref{subsec:wigglyPermutations}.

\begin{proposition}
The Hasse diagram of the wiggly lattice~$\wigglyLattice_n$ is isomorphic to the graph of the wigglyhedron~$\wigglyhedron_n$ oriented in the direction
\(
\b{\omega} \eqdef \sum_{i \in [n]} 4ni (\b{e}_{2i-1} + \b{e}_{2i}) + \sum_{j \in [2n]} j \b{e}_j.
\)
\end{proposition}

\begin{proof}
Consider an increasing flip~$T \to T'$ between two wiggly pseudotriangulations~$T$ and~$T'$ with~$T \ssm \{\alpha\} = T' \ssm \{\alpha'\}$.
Denote by~$u$ and~$v$ the labels incident to~$\alpha$ and respectively below and above~$\alpha$ in the corner labeling of~$T$ described in \cref{def:bijection1}, and let~$\bar u \eqdef \lceil u/2 \rceil$ and~$\bar v \eqdef \lceil v/2 \rceil$.
We have~$u < v$ because the flip~$T \to T'$ is increasing, so that~$\bar u \le \bar v$.

Observe now that~${\b{p}(T') - \b{p}(T)}$ is orthogonal to the $(2n-2)$-dimensional cone of the wiggly fan generated by the $\b{g}$-vectors of the wiggly arcs in~$T \cap T'$, and points from the side containing~$\b{g}(\alpha)$ to the side containing~$\b{g}(\alpha')$.
Hence, $\b{p}(T') - \b{p}(T)$ is a positive multiple of~$-\b{c}(\alpha, T)$.
From \cref{def:cvectors}, we observe that
\begin{itemize}
\item $\dotprod{\b{c}(\alpha, T)}{\b{e}_{2i-1} + \b{e}_{2i}} = 1/2$ if~$i = \bar u \ne \bar v$, $-1/2$ if~$i = \bar v \ne \bar u$, and $0$ otherwise,
\item $|\dotprod{\b{c}(\alpha, T)}{\b{e}_j}| \le 1/2$ for all~$j \in [2n]$.
\end{itemize}
We now distinguish two situations.
If~$\bar u \ne \bar v$, then
\[
\dotprod{-\b{c}(\alpha, T)}{\b{\omega}} = \sum_{i \in [n]} 4ni \dotprod{-\b{c}(\alpha, T)}{\b{e}_{2i-1} + \b{e}_{2i}} + \sum_{j \in [2n]} j \dotprod{-\b{c}(\alpha, T)}{\b{e}_j} \ge 2n(\bar v - \bar u) - n > 0.
\]
If~$\bar u = \bar v$, then~$u = 2k-1$ and~$v = 2k$ for some~$k \in [n]$, and we have
\[
\dotprod{\b{c}(\alpha, T)}{\b{\omega}} = \!\! \sum_{i \in [n]} 4ni \dotprod{-\b{c}(\alpha, T)}{\b{e}_{2i-1} + \b{e}_{2i}} + \!\! \sum_{j \in [2n]} j \dotprod{-\b{c}(\alpha, T)}{\b{e}_j} = \frac{2k}{2}-\frac{2k-1}{2} = 1/2 > 0.
\]
We conclude in both cases that~$\dotprod{\b{c}(\alpha, T)}{\b{\omega}} > 0$, so that we recover the increasing flip graph by orienting the graph of the wigglyhedron~$\wigglyhedron_n$ in direction~$\b{\omega}$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Cambrian considerations}

In this section, we fix a signature~$\delta \in \{+,-\}^n$. By convention, we set~$\delta_0 = \delta_{n+1} = 0$.
We connect the $\delta$-Cambrian lattice~\cite{Reading-CambrianLattices} to the wiggly lattice of \cref{sec:combinatorics} and the $\delta$-associahedron of~\cite{HohlwegLange} to the wigglyhedron of \cref{sec:geometry}.
\cref{def:CambrianTriangulations,def:CambrianPermutations,def:CambrianWigglyPseudotriangulations,def:CambrianWigglyPermutations,prop:CambrianLattice} are illustrated in \cref{fig:wigglyCambrian}.

\begin{figure}
\centerline{\includegraphics[scale=1.1]{wigglyCambrian1}}
\vspace{-.3cm}
\centerline{\includegraphics[scale=1.1]{wigglyCambrian2}}
\vspace{-.5cm}
\centerline{\includegraphics[scale=1.1]{wigglyCambrian3}}
\caption{Cambrian lattices as intervals of wiggly lattices.}
\label{fig:wigglyCambrian}
\end{figure}

\begin{definition}
\label{def:CambrianTriangulations}
The \defn{$\delta$-gon} is a convex $(n+2)$-gon with a point with abscissa~$i$ and with ordinate of the same sign as~$\delta_i$  for all~$0 \le i \le n+1$.
A \defn{$\delta$-triangulation} is a triangulation of the $\delta$-gon.
The \defn{$\delta$-triangulation lattice} is the transitive closure of the slope increasing flip graph on $\delta$-triangulations.
%If~$T$ and~$T'$ are two $\delta$-triangulations with~$T \ssm \{d\} = T' \ssm \{d'\}$, then the flip from~$T$ to~$T'$ is \defn{increasing} if the slope of~$d$ is smaller than the slope of~$d'$.
\end{definition}

\begin{definition}
\label{def:CambrianPermutations}
A \defn{$\delta$-permutation} is a permutation of~$[n]$ which avoids~$\cdots ik \cdots j \cdots$ with~$i < j < k$ and~$\delta_j = -$, and~$\cdots j \cdots ki \cdots$ with~$i < j < k$ and~$\delta_j = +$.
The \defn{$\delta$-permutation lattice} is the sublattice of the weak order induced by $\delta$-permutations.
\end{definition}

\begin{definition}
\label{def:CambrianWigglyPseudotriangulations}
The \defn{$\delta$-wiggly arcs} are the arcs~$(0, j, {[1,j[}, \varnothing)$ for~$\delta_j = {+}$ and~$(0, j, \varnothing, {[1,j[})$ for~$\delta_j = {-}$.
A \defn{$\delta$-wiggly pseudotriangulation} is a wiggly pseudotriangulation containing all $\delta$-wiggly arcs.
The \defn{$\delta$-wiggly pseudotriangulation lattice} is the transitive closure of the increasing flip graph on $\delta$-wiggly pseudotriangulations.
\end{definition}

\begin{definition}
\label{def:CambrianWigglyPermutations}
A \defn{$\delta$-wiggly permutation} is a wiggly permutation~$\sigma$ of~$[2n]$ such that
% for all~$j \in [n]$, we have~ and~$\sigma^{-1}(i) < \sigma^{-1}(2j-1)$ if~$\delta_j > 0$ and~$\sigma^{-1}(i) > \sigma^{-1}(2j)$ if~$\delta_j < 0$ for all~$i \in [2j-2]$.
%\begin{itemize}
%\item $\sigma^{-1}(2j-1) > \sigma^{-1}(2j)$ for all~$j \in [n]$,
%\item $\sigma^{-1}(i) < \sigma^{-1}(2j-1)$ if~$\delta_j > 0$ and~$\sigma^{-1}(i) > \sigma^{-1}(2j)$ if~$\delta_j < 0$ for all~$j \in [n]$ and~$i \in [2j-2]$.
%\end{itemize}
\[
\delta_j = {+} \quad \Longrightarrow \quad \sigma^{-1}(i) \le \sigma^{-1}(2j-1)
\qquad\text{and}\qquad
\delta_j = {-} \quad \Longrightarrow \quad \sigma^{-1}(2j) \le \sigma^{-1}(i)
\]
%$\delta_j > 0 \implies \sigma^{-1}(i) \le \sigma^{-1}(2j-1)$ and~$\delta_j < 0 \implies \sigma^{-1}(i) \ge \sigma^{-1}(2j)$
for all~$1 \le i \le 2j \le 2n$.
The \defn{$\delta$-wiggly permutation lattice} is the interval of the wiggly permutation lattice~$\wigglyLattice_n$ given by $\delta$-wiggly permutations.
\end{definition}

\begin{remark}
We had to make a choice in \cref{def:CambrianWigglyPseudotriangulations,def:CambrianWigglyPermutations}.
For~$\delta_j = {+}$, we could replace ``$(0, j, {[1,j[}, \varnothing)$'' by~``$(j, n+1, {]j,n]}, \varnothing)$'' in \cref{def:CambrianWigglyPseudotriangulations}, and ``$\sigma^{-1}(i) \le \sigma^{-1}(2j-1)$ for~$1 \le i \le 2j$'' by ``$\sigma^{-1}(i) \le \sigma^{-1}(2j-1)$ for~$2j-1 \le i \le 2n$'' in \cref{def:CambrianWigglyPermutations}.
Independently, for~$\delta_j = {-}$, we could replace ``$(0, j, \varnothing, {[1,j[})$'' by~``$(j, n+1, \varnothing, {]j,n]})$'' in \cref{def:CambrianWigglyPseudotriangulations}, and ``$\sigma^{-1}(2j) \le \sigma^{-1}(i)$ for~$1 \le i \le 2j$'' by ``$\sigma^{-1}(2j-1) \le \sigma^{-1}(i)$ for~$2j-1 \le i \le 2n$'' in \cref{def:CambrianWigglyPermutations}.
\cref{prop:CambrianLattice,prop:CambrianFan} would still hold.
\end{remark}

\begin{proposition}
\label{prop:CambrianLattice}
For any signature~$\delta \in \{+,-\}^n$, the $\delta$-triangulation lattice, $\delta$-permutation lattice, \mbox{$\delta$-wiggly} pseudotriangulation lattice, and $\delta$-wiggly permutation lattice are all isomorphic, and are known as the \defn{$\delta$-Cambrian lattice}~\cite{Reading-CambrianLattices}.
\end{proposition}

\begin{proof}
We only describe the following bijections
\begin{center}
\begin{tikzcd}[column sep=3cm]
\text{$\delta$-triangulations} \arrow[d] & \text{$\delta$-permutations} \arrow[l] \\
\text{$\delta$-wiggly pseudotriangulations} \arrow[r] & \text{$\delta$-wiggly permutations} \arrow[u]
%\text{$\delta$-triangulations} \arrow[d, "\lambda"'] & \text{$\delta$-permutations} \arrow[l, "\zeta"'] \\
%\text{$\delta$-wiggly pseudotriangulations} \arrow[r, "\mu"'] & \text{$\delta$-wiggly permutations} \arrow[u, "\nu"']
\end{tikzcd}
\end{center}
illustrated in \cref{fig:wigglyCambrian}.
We let the readers convince themselves that they are well-defined, that they compose to the identity, and that they preserve the flip structure. (We also invite the reader to work out the explicit description of the 8 missing arrows.)
For~$I \subseteq [n]$, we set~$I^+ \eqdef \set{i \in I}{\delta_i = {+}}$ and~$I^- \eqdef \set{i \in I}{\delta_i = {-}}$.

\para{From $\delta$-triangulations to $\delta$-wiggly pseudotriangulations}
Replace each diagonal~$(i,j)$ of the $\delta$-gon by the wiggly arc~$(0, j, [i] \cup {]i,j[}^-, {]i,j[}^+)$ if~$\delta_i \ge 0$ and~$(0, j, {]i,j[}^-, [i] \cup {]i,j[}^+)$ if~$\delta_i \le 0$.

\para{From $\delta$-wiggly pseudotriangulations to $\delta$-wiggly permutations}
Specialization of \cref{def:bijection1}.

\para{From $\delta$-wiggly permutations to $\delta$-permutations}
For all~$j \in [n]$, erase~$2j-1$ and replace~$2j$ by~$j$ if~$\delta_j = {+}$, and erase~$2j$ and replace~$2j-1$ by~$j$ if~$\delta_j = {-}$.

\para{From $\delta$-permutations to $\delta$-triangulations}
Already described in~\cite{Reading-CambrianLattices}.
Given a permutation~$\pi$, construct the union of the upper hulls of the point sets~$[n]^- \symdif \pi([j])$ for all~$j \in \{0, \dots, n\}$.
\end{proof}

Recall that the $\delta$-Cambrian lattice can be realized by the $\delta$-Cambrian fan~\cite{ReadingSpeyer} and the $\delta$-associahedron of~\cite{HohlwegLange}.
Namely, to each diagonal~$(i,j)$ of the $\delta$-gon, we associate the subset~$X(\delta)$ of points of the $\delta$-gon distinct from~$0$ and~$n+1$, and located below the line~$(i,j)$, including~$i$ if and only if~$\delta_i = {+}$ and~$j$ if and only if~$\delta_j = {+}$.
The \defn{$\delta$-Cambrian fan} is the fan with a ray~$\b{g}(d) \eqdef \pi(\one_{X(d)})$ for each internal diagonal~$d$ of the $\delta$-gon and a cone~$\R_{\ge0} \setangle{\b{g}(d)}{d \in D^\circ}$ for each dissection~$D$ of the $\delta$-gon.
The \defn{$\delta$-associahedron} is the polytope defined by the equality~$\dotprod{\one_n}{\b{x}} = \binom{n+1}{2}$ and the inequalities~$\dotprod{\b{g}(d)}{\b{x}} \le \binom{|X(d)|+1}{2}$ for all internal diagonals~$d$ of the $\delta$-gon.

\begin{proposition}
\label{prop:CambrianFan}
The $\delta$-associahedron~$\Asso_\delta$ is normally equivalent to the face of the wigglyhedron~$\wigglyhedron_n$ corresponding to the wiggly pseudodissection formed by the $\delta$-wiggly arcs.
\end{proposition}

\begin{proof}
Following the proof of \cref{prop:CambrianLattice}, consider the map~$\zeta$ sending a diagonal~$(i,j)$ of the $\delta$-gon to the wiggly arc $(0, j, [i] \cup {]i,j[}^-, {]i,j[}^+)$ if~$\delta_i \ge 0$ and~$(0, j, {]i,j[}^-, [i] \cup {]i,j[}^+)$ if~$\delta_i \le 0$.
Let~$a \eqdef (i,j)$ and~$a' \eqdef (i',j')$ be two crossing diagonals of the~$\delta$-gon.
Then any $\delta$-triangulations~$T$ and~$T'$ with~$T \ssm \{a\} = T' \ssm \{a'\}$ contain the diagonals~$b \eqdef (i',j)$ and~$b' \eqdef (i,j')$ of the $\delta$-gon, and~$\b{g}(a) + \b{g}(a') = \b{g}(b) + \b{g}(b')$.
Moreover, if~$\alpha \eqdef \zeta(a)$, $\alpha' \eqdef \zeta(a')$, $\beta \eqdef \zeta(b)$ and~$\beta' \eqdef \zeta(b')$, then~$\b{g}(\alpha) + \b{g}(\alpha') = \b{g}(\beta) + \b{g}(\beta')$.
We thus conclude that $\zeta$ preserves the linear dependences among $\b{g}$-vectors.
Hence, there is a linear map~$Z : \HH_n \to \HH_{2n}$ such that~$Z \big( \b{g}(d) \big) = \b{g} \big( \zeta(d) \big)$ for all diagonals~$d$ of the $\delta$-gon.
This map is a normal equivalence from $\delta$-associahedron~$\Asso_\delta$ to the face of the wigglyhedron~$\wigglyhedron_n$ corresponding to the wiggly pseudodissection formed by the $\delta$-wiggly~arcs.
\end{proof}

%\vincent{todo: find an explicit affine map between the Hohlweg-Lange $\delta$-associahedron and the face of the wigglyhedron corresponding to all $\delta$-wiggly arcs.}

\begin{remark}
In their work on polytopal realizations of $\b{g}$-vector fans of finite type cluster algebras~\cite{HohlwegPilaudStella}, C.~Hohlweg, V.~Pilaud and S.~Stella already constructed a \defn{universal associahedron}~$\polytope{U}_n$ which somehow contains all Cambrian fans (actually all $\b{g}$-vector fans from any initial cluster seed).
More precisely, the $\delta$-Cambrian fan for each~$\delta \in \{+,-\}^n$  appears as a section of the normal fan of the universal associahedron~$\polytope{U}_n$, which is a $(n^2+n-4)/2$-dimensional polytope.
In contrast, \cref{prop:CambrianFan} states that the $\delta$-Cambrian fan for each~$\delta \in \{+,-\}^n$ appears as a link of the normal fan of the wigglyhedron~$\wigglyhedron_n$, which is a $(2n-1)$-dimensional polytope.
There seem to be no connection between these two universal polytopes.
However, this certainly raises the question to construct analogues of the wigglyhedron for all finite Coxeter groups.
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Wiggly pseudotriangulations of planar point sets}

We now consider \cref{thm:wigglyhedron,conj:Hamiltonian} for wiggly pseudotriangulations of arbitrary point sets in the plane (neither necessarily aligned, nor necessarily in general position).

\begin{definition}
Fix a point set~$P$ of the plane, and an arbitrary total order~$<$ on~$P$.
A \defn{wiggly arc} is a quadruple~$(p,q,R,S)$ where~$p < q \in P$ and the sets~$R$ and~$S$ form a partition of the points of~$P$ located in the open segment joining~$p$ to~$q$.
Two wiggly arcs~$(p,q,R,S)$ and~$(p',q',R',S')$ are \defn{crossing} if 
\begin{itemize}
\item either the segments~$[p,q]$ and~$[p',q']$ cross, 
\item or~$(R \cap S') \cup (\{p,q\} \cap S') \cup (R \cap \{p',q'\}) \ne \varnothing \ne (R' \cap S) \cup (\{p',q'\} \cap S) \cup (R' \cap \{p,q\})$,
\end{itemize}
A set~$X$ of wiggly arcs is \defn{pointed} if for any~$p \in P$, the wiggly arcs of~$X$ with an endpoint at~$p$ generate a pointed cone.
The \defn{wiggly complex}~$\wigglyComplex_P$ is the simplicial complex of pairwise pointed and non-crossing subsets of wiggly arcs.
Note that the boundary wiggly arcs are irrelevant, which allows us to consider a reduced wiggly complex~$\wigglyComplex_P$ induced by internal wiggly arcs.
A \defn{wiggly pseudotriangulation} of~$P$ is a facet of~$\wigglyComplex_P$.
The \defn{wiggly flip graph}~$\wigglyFlipGraph_P$ is the adjacency graph of the facets of~$\wigglyComplex_P$.
Note that, by construction, $\wigglyFlipGraph_P$ is regular.
See \cref{fig:wigglyComplexSquarre} for an illustration.
%
\begin{figure}[!h]
\centerline{\includegraphics[scale=1]{wigglyComplexSquare}}
\caption{The wiggly complex~$\wigglyComplex_P$ (left) and the wiggly flip graph~$\wigglyFlipGraph_P$ (right) of a point set~$P$.}
\label{fig:wigglyComplexSquarre}
\end{figure}
\end{definition}

\begin{conjecture}
\label{conj:polytopality}
For any point set~$P$ in the plane, the wiggly complex~$\wigglyComplex_P$ is the boundary complex of a simplicial polytope.
\end{conjecture}

Note that in \cref{conj:polytopality}, the case of aligned points is given by the wigglyhedron of \cref{thm:wigglyhedron}, while the case of points in general position is given by the pseudotriangulation polytope of~\cite{RoteSantosStreinu-polytope} (see also \cite{RoteSantosStreinu-pseudotriangulations} for a nice survey on pseudotriangulations).

\begin{conjecture}
\label{conj:Hamiltonian2}
For any point set~$P$ in the plane, the wiggly flip graph~$\wigglyFlipGraph_P$ is Hamiltonian.
\end{conjecture}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Further questions}

A few additional questions to arouse your curiosity:
\begin{enumerate}
\item Can the construction of \cref{thm:wigglyhedron} be adapted to provide a more combinatorial construction of the polytope of pseudotriangulations~\cite{RoteSantosStreinu-polytope}, that would only depend on the order type (aka oriented matroid~\cite{BjornerLasVergnasSturmfelsWhiteZiegler}) of the point configuration?
\item What is the dual interpretation of wiggly pseudotriangulations as pseudoline arrangements? (see~\cite{PilaudPocchiola} for context) Can this interpretation be extended to other finite Coxeter groups? (see~\cite{CeballosLabbeStump} for context).
\item What is the multi wiggly complex?
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Categorical interpretation of wiggly pseudotriangulations}
The aim of this section is to explain a categorical interpretation of wiggly arcs.
We recall results of Khovanov--Seidel~\cite{kho.sei:02} that realise wiggly arcs as objects of a triangulated category, such that intersections between wiggly arcs correspond to morphism in this category.
We also discuss categorical conditions for two wiggly arcs to be compatible as in~\cref{def:compatible}.

% Curves up to isotopy in a disk with marked points constitute a geometric model of certain objects in a triangulated category.
% This correspondence, due to Khovanov--Seidel~\cite{kho.sei:02}, and will enable us to build categorical interpretations of the combinatorial results obtained in this paper.
% The aim of this section is to give a concise account of this correspondence and some of its useful features; we refer the reader to~\cite{kho.sei:02} for details.

\subsection{A triangulated category associated to a quiver}
We begin by describing the category of interest, which is a subcategory of \(D^b(A_m\text{-mod})\) for a certain finite-dimensional algebra \(A_m\).
That is, a subcategory of the bounded derived category of (finitely generated) \(A_m\)-modules.
The algebra \(A_m\), called the \emph{zigzag algebra}, is defined as a quotient of the path algebra of the quiver \(Q_m\) with \(m\) vertices depicted in~\cref{fig:am-quiver}.
\begin{figure}[h]
  \centering
  \begin{tikzcd}
    1 \arrow[bend left]{r} & 2 \arrow[bend left]{r}\arrow[bend left]{l}& \cdots \arrow[bend left]{r}\arrow[bend left]{l}& m \arrow[bend left]{l}
  \end{tikzcd}
  \caption{The quiver \(Q_m\). The algebra \(A_m\) is a quotient of the path algebra \(kQ_m\).}
  \label{fig:am-quiver}
\end{figure}
Let \(kQ_m\) be the path algebra of \(Q_m\), graded by path length.
Let \(e_i\) denote the idempotent path at the vertex \(i\), and let \((i|j)\) denote the arrow from vertex \(i\) to a (neighbouring) vertex \(j\).
A longer path, which is a composition of arrows, will be denoted by \((i_1|i_2|\cdots|i_n)\).

\begin{definition}
  The zigzag algebra \(A_m\) corresponding to the quiver \(Q_m\) depicted in~\cref{fig:am-quiver} is the quotient of \(kQ_m\) by the following relations:
  \begin{itemize}
  \item \((i|j|l) = 0\) whenever \(i \neq l\);
  \item \((i|j|i) = (i|l|i)\) for every \(j\) and \(l\) such that there are arrows between \(i\) and \(j\) and between \(i\) and \(l\).
  \end{itemize}
\end{definition}

\begin{remark}
  The definition of the zigzag algebra as used here originally appeared in~\cite[Section 3]{hue.kho:01}.
  This is an ``un-signed'' version of the zigzag algebra.
  It is technically better to use a ``signed'' version of the zigzag algebra, and the comparison between the two is discussed in~\cite[Remark 6.6]{bap.deo.lic:20}.
  As the two versions are isomorphic in our example, we ignore this technicality and use the easier definition.
\end{remark}

Let \(P_i\) denote the indecomposable projective \(A_m\)-module defined as \(P_i = A_m\cdot e_i\).
For any integer \(j\), let \(P_i \langle j \rangle\) denote its graded shift by \(j\).
Let \(\mathcal{C}_m\) be the smallest triangulated subcategory of \(D^b(A_m\text{-mod})\) containing all the objects \(P_i\langle j \rangle\) (see, e.g.,~\cite[Section 2.3.3]{bap.deo.lic:22} for further discussion on the construction).

The category \(\mathcal{C}_m\) is \(k\)-linear and triangulated, where \([1]\) denotes the triangulated shift.
For \(X, Y \in \mathcal{C}_m\), let \(\Hom^{g,h}(X,Y)\) denote the space of morphisms in \(\mathcal{C}_m\) of internal degree \(g\) and homological degree \(h\).
In other words, we have
\[\Hom^{g,h}(X,Y) = \Hom(X,Y\langle g \rangle[h]).\]
Furthermore, for any integers \(a\) and \(b\), we have
\[\Hom^{g,h}(X\langle g' \rangle[h'], Y\langle g' \rangle[h']) = \Hom^{g,h}(X,Y).\]

\begin{remark}\label{rem:generating-morphisms}
  It is easy to check that the generating objects \(P_i\) (and hence their graded and homological shifts) satisfy the following:
  \begin{itemize}
  \item \(\Hom^{g,h}(P_i,P_i) = 0\) if \(g \neq 0, 2\) and \(h \neq 0\);
  \item \(\Hom^{g,0}(P_i,P_i) = k\) if \(g = 0, 2\);
  \item \(\Hom^{g,h}(P_i, P_{i\pm1}) = 0\) if \(g \neq 1\) and \(h \neq 0\);
  \item \(\Hom^{1,0}(P_i, P_{i\pm1}) = k\), and
  \item \(\Hom^{g,h}(P_i, P_j) = 0 \) otherwise.
  \end{itemize}
\end{remark}

We now describe the full subcategory of \emph{linear complexes} of \(\mathcal{C}_m\).
It is convenient to think of an object \(P^{\bullet}\) of \(\mathcal{C}_m\)
as bigraded (projective) \(A_m\)-module equipped with a differential \(d \colon P^{\bullet} \to P^{\bullet}\).
More precisely, we can write
\[P^{\bullet} = \bigoplus M_i\langle g \rangle[h],\]
where \(g\) is the internal grading shift, and \(h\) is the homological grading shift, and 
each object \(M_i\) is a direct sum of copies of objects in \(\{P_1,\ldots, P_m\}\).

Recall that an object of the form \(M_i[h]\) is placed in homological degree \(-h\);
since the differential \(d\) raises the homological degree by one, it \emph{lowers} the homological shift by one.
That is, we have
\[d \colon M_i\langle g \rangle[h] \to \bigoplus M_{i'}\langle g' \rangle[h-1].\]
In other words, \(d\) maps \(P^h\) to \(P^{h-1}\).
\begin{definition}\label{def:linear-complex}
  Let \(P^{\bullet}\) be an object of \(\mathcal{C}_m\), described as a bigraded module
  \[P^{\bullet} = \bigoplus M_i\langle g \rangle[h]\]
  together with an internal differential \(d \colon P^{\bullet} \to P^{\bullet + 1}\).
  We say that \(P^{\bullet}\) has \emph{level \(\ell\)} for some \(\ell \in \mathbb{Z}\) if for each bigraded piece \(M_i \langle g \rangle[h]\) of \(P^{\bullet}\), we have \(g + h = \ell\).
  We say that \(P^{\bullet}\) is \emph{linear} if \(P^{\bullet}\) has level \(\ell = 0\).

\end{definition}
Let \(\mathcal{A}_m\) be the smallest full subcategory of \(\mathcal{C}_m\) that is closed under isomorphism, which contains all the linear complexes.
\begin{proposition}\label{prop:standard-heart}
  The category \(\mathcal{A}_m\) is abelian.
  Furthermore, it is the heart of a bounded \(t\)-structure on \(\mathcal{C}_m\), which we call the \emph{standard \(t\)-structure} on \(\mathcal{C}_m\).
\end{proposition}
\asilata{Find a reference for this and possibly a cleaner statement.}

Also recall the following statements for future use.
\begin{defprop}\label{defprop:cohomology}
  Let \((\mathcal{C}^{\le 0}, \mathcal{C}^{\geq 0})\) be a bounded \(t\)-structure on a triangulated category \(\mathcal{C}\) with heart \(\mathcal{A} = \mathcal{C}^{\leq 0} \cap \mathcal{C}^{\geq 0}\).
  For each integer \(i\), we have the \(i\)th truncation functors
  \[\tau^{\leq i} \colon \mathcal{C} \to \mathcal{C}^{\leq i}\]
  and
  \[\tau^{\geq i} \colon \mathcal{C} \to \mathcal{C}^{\geq i},\]
  which are left and right adjoint respectively\asilata{Check left vs right.} to the inclusion functors \(\mathcal{C}^{\leq 0} \to \mathcal{C}\) and \(\mathcal{C}^{\geq 0} \to \mathcal{C}\).
  The \(i\)th cohomology functor is defined to be the composition \(\tau^{\leq i} \circ \tau^{\geq i}\).
  It is equal to the composition \(\tau^{\geq i} \circ \tau^{\leq i}\), and maps an object of \(\mathcal{C}\) to an object of \(\mathcal{A}\).
\end{defprop}
In the presence of a bounded \(t\)-structure, every object of \(\mathcal{C}\) has a canonical filtration whose factors are shifts of its cohomology objects.
\begin{defprop}\label{defprop:cohomology-filtration}
  Let \(\mathcal{C}\) be any triangulated category, and \(\mathcal{A} \subset \mathcal{C}\) the heart of a bounded \(t\)-structure on \(\mathcal{C}\).
  Let \(X\) be a non-zero object of \(\mathcal{C}\), and let \(k\) and \(l\) be the maximum and minimum integers respectively for which \(H^k(X)\) and \(H^{l}(X)\) are non-zero.
  \asilata{Find references for cohomology and cohomology filtration.}
  Then \(X\) has a unique filtration of the form
  \[
    \begin{tikzcd}[column sep=0.1cm]
      0  \arrow{rr} && X_{-l} \arrow{dl} \arrow{rr} && \cdots \arrow{rr} \arrow{dl} && X_{-k+1} \arrow{rr} \arrow{dl} && X_{-k} = X \arrow{dl}\\
      & H^l(X)[-l] \arrow[dashed]{ul}{+1} && H^{l+1}(X)[-l-1] \arrow[dashed]{ul}{+1} && \cdots \arrow[dashed]{ul} &&H^k(X)[-k] \arrow[dashed]{ul}{+1}
    \end{tikzcd}
  \]  
\end{defprop}

\subsection{The Khovanov--Seidel curve model}
In this section we briefly recall how to recover an object of \(\mathcal{C}_{n+1}\) from a wiggly arc on \(n+2\) points on a line.
In fact, the procedure works on more general arcs as well --- namely, those that are not necessarily \(x\)-monotone.
All the results of this section are due to Khovanov and Seidel~\cite{kho.sei:02}.

Consider \(n+2\) points arranged in a straight horizontal line, and labelled \(0,1, \ldots, n+1\) as in~\cref{sec:combinatorics}.
In addition to the wiggly arcs, consider isotopy classes of curves whose endpoints lie on distinct marked points, and which do not cross themselves or any marked point in their interior.
For convenience, we refer to such an isotopy class as an \emph{arc}.

The Khovanov--Seidel correspondence (explained in Section 4 of~\cite{kho.sei:02}) associates objects of \(\mathcal{C}_{n+1}\) (up to grading\asilata{Should this say homological?} shift) to curves on a configuration of \(n+2\) marked points arranged in a straight line.

In particular, the line segment joining point \(i\) to point \(i+1\) is sent to the object \(P_{i+1}\) in the category \(\mathcal{C}_{n+1}\).
It is easy to check that the (minimal) intersection numbers between any two such segments counts the number of morphisms\asilata{Discuss (or choose to ignore) grading.} between the associated objects.
In particular, the self-intersection number of a single segment is computed to be \(2\), as the second copy of the segment can be slightly deformed to be disjoint from the first copy except at the two endpoints.
In fact, it is true for any pair of curves that its minimal intersection number is the number of morphisms between the associated objects (\cite[Lemma 4.11]{kho.sei:02}).

\begin{lemma}
  Every wiggly arc corresponds to an object in the heart \(\mathcal{A}_m\) of the standard \(t\)-structure on \(\mathcal{C}_m\).
\end{lemma}
\begin{proof}
  Let \(P^{\bullet}\) be the complex built from an arbitrary (not necessarily wiggly) arc.
  The procedure to build up the complex of projective modules from a curve explicitly describes the differentials.

  Each component of these differentials is a map of the form
  \[P_i\langle g \rangle[h] \to P_j\langle g' \rangle[h-1].\]
  First consider a component of the differential from some \(P_i\langle g \rangle\) to some \(P_j\langle g' \rangle\) for \(i \neq j\).
  We see from~\cref{rem:generating-morphisms} that any such map must take \(P_i\langle g \rangle\) to \(P_j \langle g+1 \rangle\).
  In other words, the differentials has the form
  \[P_i\langle g \rangle[h] \to P_j\langle g+1 \rangle[h-1]\]
  if \(i \neq j\).
  Since \(P_i\langle g \rangle[h]\) and \(P_j\langle g+1 \rangle[h-1]\) have the same level, any differential from a graded shift of \(P_i\) to a graded shift of \(P_j\) keeps the level unchanged.
  The other possibility is a map from \(P_i\langle g \rangle\) to \(P_i\langle g' \rangle\).
  Once again, it is clear from~\cref{rem:generating-morphisms} that such differentials have the form
  \[P_i \langle g \rangle[h] \to P_i\langle g' \rangle[h-1],\]
  where \(g'\) is either \(g\) or \(g + 2\).
  Any component of the form
  \[P_i \langle g \rangle[h] \to P_i\langle g \rangle[h-1]\]
  is necessarily a multiple of the identity map, and identity maps do not appear in the Khovanov--Seidel complex.
  Hence the only possibility is multiples of the loop map, which have the form
  \[P_i \langle g \rangle[h] \to P_i\langle g+2 \rangle[h-1].\]  
  Therefore any portion of the differential from a graded shift of a \(P_i\) to another graded shift of itself raises the level by one.

  Computing the \(j\)th cohomology of \(P^{\bullet}\) simply amounts to truncating the complex to level \(j\).
  The object \(P^{\bullet}\) is in the heart if and only if it is concentrated in a single cohomological degree, which means that its differential has no components of the form
  \[P_i \langle g \rangle[h] \to P_i\langle g+2 \rangle[h-1].\]
  This is the case if and only if the corresponding curve does not loop back around any point; that is, if and only if it is \(x\)-monotone.
  The lemma is proved.
\end{proof}

\subsection{Cohomology filtrations of spherical objects}
In this section we explain how wiggly pseudotriangulations govern the combinatorial structure of a natural categorical decomposition of spherical objects of \(\mathcal{C}_m\).

Recall from~\cref{defprop:cohomology-filtration} that any object \(X\) of \(\mathcal{C}_m\) has a canonical cohomology filtration whose factors are homological shifts of objects in \(\mathcal{A}_m\).
For general \(X\), the cohomology filtration can be arbitrary.
However, it turns out that the cohomology filtration is highly constrained if \(X\) is spherical, and these constraints are given by wiggly pseudotriangulations, as explained in the following proposition.
\begin{proposition}
  Let \(X \in \mathcal{C}_{n+1}\) be a spherical object.
  Consider its cohomology filtration
  \[
    \begin{tikzcd}[column sep=0.3cm]
      0 \arrow{rr} && X_{-l} \arrow{dl} \arrow{rr} && \cdots \arrow{rr} \arrow{dl} && X_{-k} = X \arrow{dl}\\
      & H^l(X)[-l] \arrow[dashed]{ul}{+1} && H^{l+1}(X)[-l-1] \arrow[dashed]{ul}{+1} & \cdots & H^k(X)[-k] \arrow[dashed]{ul}{+1}
    \end{tikzcd}
  \]
  Consider the set of all nonzero objects \(\{H^j(X)\}\) taken without multiplicity, and let \(T\) be corresponding set of wiggly curves.
  We have the following results.
  \begin{enumerate}
  \item 
    \(T\) is a subset of a wiggly pseudotriangulation.
  \item \(T\) contains at most one of the two irrelevant arcs.
  \end{enumerate}
\end{proposition}
\begin{proof}
  The proof of this proposition is an easy consequence of analysing cohomology filtration from the point of view of arcs on the disk with marked points.
  \asilata{TODO.}
\end{proof}

\subsection{Categorical constraints on compatibility of curves}
In this section, we give a categorical criterion for two wiggly arcs to be compatible as in~\cref{def:compatible}.
As such, this section is somewhat dual to the previous section --- our criterion determines whether a pair of wiggly arcs can belong together in a single wiggly pseudotriangulation, whereas the previous section generated a wiggly pseudotriangulation given any object of the category.
\begin{proposition}
  Consider two wiggly arcs \(\alpha\) and \(\alpha'\), and let \(X\) and \(X'\) be objects in the standard heart \(\mathcal{A}_m\) corresponding to \(\alpha\) and \(\alpha'\) respectively.
  Then \(\alpha\) and \(\alpha'\) are compatible in the sense of~\cref{def:compatible} if and only if \(\Ext^1(X\langle g \rangle, X\langle g' \rangle) = 0\) for any two integers \(g\) and \(g'\).
\end{proposition}
\begin{remark}
  This proposition is similar to the condition of being compatible in a cluster-tilting object.
  \asilata{Expand and make precise.}
\end{remark}
\begin{remark}
  As stated, this proposition is not expected to be true in greater generality.
  However, we expect to see similar constraints on compatibility, which we hope to turn to in forthcoming work.
  \asilata{Decide what to say here, because the forthcoming work is far from complete, although I do have counterexamples in type \(D_4\). Discussing how to interpret these counterexamples may not involve too much more work?}
\end{remark}
\begin{question}
  How to best relate this section to the previous one purely categorically?
  That is, given the categorical constraint on compatibility of arcs, is it also clear that it gives a constraint on objects appearing in the cohomology filtration?
  \asilata{Make more precise.}
\end{question}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgements}

This work started during the \emph{Winterbraids XIII: School on low dimensional topology, braid groups and interactions}. We are grateful to the organizers Bérénice Delcroix-Oger, Peter Feller, Vincent Florens, Hoel Queffelec for this fruitful event.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\clearpage
\addtocontents{toc}{ \vspace{.1cm} }
\bibliographystyle{alpha}
\bibliography{wigglyhedra}
\label{sec:biblio}

\end{document}

%%% Local variables:
%%% TeX-command-extra-options: "-pvc"
%%% End: