### 2024-10-08 exposeBerlin2410/exposeBerlin2410.tex
start presentation Berlin

\documentclass[12pt,titlepage,landscape,a4paper]{article}

%QUELQUES PACKAGES PLUS OU MOINS UTILES
\special{landscape}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsfonts, amsmath, amssymb, amsthm, mathtools}
\usepackage{aeguill}
\usepackage{graphics, graphicx}
\usepackage{xcolor}
\usepackage{geometry}
\usepackage{paralist}
\usepackage{multido,ifthen}
\usepackage{tikz}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles, snakes}
\usepackage[external]{forest}
\usepackage{hyperref}
\hypersetup{colorlinks=true, citecolor=blue, linkcolor=blue, urlcolor=blue}
\usepackage{movie15}
\usepackage[abs]{overpic}
\usepackage{xargs}
\usepackage{extarrows}
\usepackage{shuffle}
\usepackage{multicol}
\usepackage{etoolbox}
\usepackage{dsfont}
\usepackage{pifont}
\usepackage{array, blkarray}
\usepackage{multicol, multirow}
\usepackage{ulem}
\usepackage{xifthen}

% ZONE DE TEXTE ET POLICE
\setlength{\topmargin}{-2.8cm}
\setlength{\textheight}{19cm}
\setlength{\oddsidemargin}{-1cm}
\setlength{\textwidth}{26.8cm}
\newcommand{\textenormal}{\fontsize{22}{28}\selectfont}
\newcommand{\textemoyen}{\fontsize{23}{27}\selectfont}
\newcommand{\textegrand}{\fontsize{55}{48}\selectfont}

% COMMANDE DE NOUVELLE PAGE
\newenvironment{slide}[1]
{
\newpage
\begin{center}
{\blue \textemoyen \uppercase{#1}}\\
\end{center}
\vspace{-1cm}
\rule{\textwidth}{0.5 pt}\\
\vspace{-.8cm}
}
{\vspace*{-3cm}}

% COMMANDE DE PARTIE
%\newcounter{partie}
%\setcounter{partie}{1}
%\newcommand{\thePartie}{Part \arabic{partie}}
\newcommand{\partie}[1]
{
\newpage
\vspace*{0pt plus 1 fill}
\begin{center}
%\thePartie \\[-.75cm]
\rule{.9\textwidth}{0.5 pt} \\[.5cm]
\fontsize{30}{30}\selectfont {\blue\uppercase{#1}} \\[-.35cm]
\rule{.9\textwidth}{0.5 pt} \\
\end{center}
\vspace*{0pt plus 1.2 fill}
%\addtocounter{partie}{1}
}

% COMMANDE D'EXPOSE
\newcommand{\expose}[2]
{
\newpage
\vspace*{0pt plus 1 fill}
\begin{center}
%\thePartie \\[-.75cm]
\rule{.8\textwidth}{0.5 pt} \\[.5cm]
\fontsize{35}{35}\selectfont {\red\uppercase{#1}} \\[-.35cm]
\rule{.8\textwidth}{0.5 pt} \\
\vspace{1cm}
\textemoyen #2
\end{center}
\vspace*{0pt plus 1.2 fill}
%\addtocounter{partie}{1}
}

% COMMANDE POUR LES BOITES
% centre
\newcommand{\cboite}[1]
{
\vspace*{.5cm}
\fcolorbox{blue}{grisclair}{
\begin{minipage}{.97\linewidth}
\vspace*{.3cm}
\begin{center} #1 \end{center}
\vspace*{-.1cm}
\end{minipage}
}
}
% alignement gauche
\newcommand{\gboite}[1]
{
\vspace*{.5cm}
\fcolorbox{blue}{grisclair}{
\begin{minipage}{.97\linewidth}
\vspace*{.3cm}
#1
\vspace{.3cm}
\end{minipage}
}
}

\newcommand{\chapitre}[1]{{\blue \fontsize{23}{25}\selectfont {Chap\,#1}}}

% AUTRES COMMANDES
% quelques couleurs manquantes
\newcommand{\orange}{\color{orange}} % couleur orange
\newcommand{\green}{\color{green}} % couleur verte
\newcommand{\violet}{\color{violet}} % couleur violet
\newcommand{\blue}{\color{blue}} % couleur bleu
\newcommand{\red}{\color{red}} % couleur rouge
\definecolor{violet}{rgb}{.5,.1,.9}
\definecolor{orange}{rgb}{.94,.57,0}
\definecolor{green}{rgb}{0.2,0.6,0.2}
\definecolor{grisclair}{gray}{1}
\definecolor{grisfonce}{gray}{.1}
\definecolor{bblue}{rgb}{.8,.8,1}
% maths
\newcommand{\set}[2]{\left\{ #1 \;\middle|\; #2 \right\}} % ensemble
\newcommand{\bigset}[2]{\big\{ #1 \;\big|\; #2 \big\}} % ensemble
\newcommand{\biggset}[2]{\bigg\{ #1 \;\bigg|\; #2 \bigg\}} % ensemble
\newcommand{\setangle}[2]{\left\langle #1 \;\middle|\; #2 \right\rangle} % ensemble
\newcommand{\dotprod}[2]{\left\langle\; #1 \;\middle|\; #2 \;\right\rangle} % produit scalaire
\newcommand{\ssm}{\smallsetminus} % small set minus
\newcommand{\symdif}{\triangle} % symmetric difference
\newcommand{\eqdef}{\mbox{\,\raisebox{0.3ex}{\normalsize\ensuremath{\mathrm:}}\ensuremath{=}\,}} % :=
\newcommand{\defeq}{\mbox{~\ensuremath{=}\raisebox{0.3ex}{\normalsize\ensuremath{\mathrm:}} }} % =:
\newcommand{\Fracfloor}[2]{\left\lfloor \frac{#1}{#2} \right\rfloor} % floor of a fraction
\newcommand{\one}{1\!\!1} % one bold
\DeclareMathOperator{\conv}{conv} % enveloppe convexe
\DeclareMathOperator{\cone}{cone} % cone engendre
\DeclareMathOperator{\vol}{vol} % volume
\newcommand{\C}{\mathbb{C}} % complexes
\newcommand{\R}{\mathbb{R}} % reels
\newcommand{\Z}{\mathbb{Z}} % entiers
\newcommand{\N}{\mathbb{N}} % naturels
\newcommand{\I}{\mathbb{I}} % set of integers
\newcommand{\K}{\mathbb{K}} % field
\newcommand{\fS}{\mathfrak{S}} % symmetric group
\newcommand{\cA}{\mathcal{A}} % algebra
\newcommand{\cF}{\mathcal{F}} % flip graph
\newcommand{\cN}{\mathcal{N}} % sorting network
\renewcommand{\b}[1]{\boldsymbol{#1}} % bold letters
\renewcommand{\c}[1]{\mathcal{#1}} % caligraphic letters
\newcommand{\f}[1]{\mathfrak{#1}} % frak letters
\newcommand{\bb}[1]{\mathbb{#1}} % bbold letters
% autres
\setlength{\parindent}{0pt} % aucune indentation dans tout le document
\graphicspath{{figures/}{figures/nodes/}} % les repertoires ou se trouvent les figures
\newcommand{\papier}[1]{{\violet\fontsize{15}{20}\selectfont #1}} % citation papier
\newcommand{\theo}[2]{\gboite{{\blue \fontsize{18}{25}\selectfont #1.} #2}}
\newcommand{\HUGE}[1]{{\fontsize{35}{33}\selectfont #1}}
\newcommand{\esperluette}{ \\ --- \& --- \\ } % esperluette stylisée nouvelle ligne
\DeclareRobustCommand{\verylongrightarrow}{\joinrel\relbar\joinrel\relbar\joinrel\relbar\joinrel\relbar\joinrel\relbar\joinrel\rightarrow}
\renewcommand{\emph}[1]{\uline{#1}}

% specific wiggly
\newcommand{\polytope}[1]{\mathds{#1}} % font polytope
\newcommand{\wigglyComplex}{\mathrm{WC}} % wiggly complex
\newcommand{\wigglyFlipGraph}{\mathrm{WFG}} % wiggly flip graph
\newcommand{\wigglyIncreasingFlipGraph}{\mathrm{WIFG}} % wiggly increasing flip graph
\newcommand{\wigglyLattice}{\mathrm{WL}} % wiggly lattice
\newcommand{\wigglyFan}{\mathrm{WF}} % wiggly fan
\newcommand{\wigglyhedron}{\polytope{W}} % wigglyhedron
\newcommand{\Asso}{\polytope{A}\mathsf{sso}} % associahedron

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\titre}{From permutahedra to associahedra, a walk through geometric and algebraic combinatorics}
\newcommand{\auteur}{Vincent Pilaud}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\fontshape{sf}\fontsize{22}{28}\selectfont % police par defaut
\sf
\pagestyle{empty} % bas de page par defaut

\vspace*{-1cm}
\begin{center}
{\blue \fontsize{60}{60}\selectfont
\uppercase{Wigglyhedra}

}

\vspace{.8cm}

A. BAPAT
(The Australian National University)

V. PILAUD
(Universitat de Barcelona)

\href{https://arxiv.org/abs/2407.11632}{\texttt{arxiv:2407.11632}}

\vspace{.2cm}
\centerline{\includegraphics[scale=1.7]{wigglyFlipGraph}\quad\includegraphics[scale=1.7]{wigglyLattice}}

\vspace{.3cm}

DMG Seminar Berlin --- October 9, 2024

\vspace*{-.6cm}

\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\partie{Triangulations \& associahedra}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\partie{Pseudotriangulations \& PPT polytope}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\partie{Wiggly pseudotriangulations \& wiggly complex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{slide}{Wiggly complex}

\emph{wiggly dissection} $=$ set of pairwise \emph{non-crossing} and \emph{pointed} wiggly arcs on $n+2$ points

\hspace{9.3cm}\includegraphics[scale=1.7]{incompatible}

\vspace{.2cm}
\emph{wiggly complex}~$\wigglyComplex_n$ $=$ simplicial complex of wiggly dissections

\vspace{.2cm}
\centerline{\includegraphics[scale=1.7]{wigglyComplex}}

%\theo{PROP}{The wiggly complex is a pure $(2n-1)$-dimensional pseudomanifold.}

\end{slide}

%%%

\begin{slide}{Wiggly complex}

\emph{wiggly dissection} $=$ set of pairwise \emph{non-crossing} and \emph{pointed} wiggly arcs on $n+2$ points

\hspace{9.3cm}\includegraphics[scale=1.7]{incompatible}

\vspace{.2cm}
\emph{wiggly complex}~$\wigglyComplex_n$ $=$ simplicial complex of wiggly dissections

\begin{align*}
f(\wigglyComplex_1) & = (1, 2) \\
f(\wigglyComplex_2) & = (1, 9, 21, 14) \\
f(\wigglyComplex_3) & = (1, 24, 154, 396, 440, 176) \\
f(\wigglyComplex_4) & = (1, 55, 729, 4002, 10930, 15684, 11312, 3232) \\
f(\wigglyComplex_5) & = (1, 118, 2868, 28110, 140782, 400374, 673274, 662668, 352728, 78384) \\[.8cm]
h(\wigglyComplex_1) & = (1, 1) \\
h(\wigglyComplex_2) & = (1, 6, 6, 1) \\
h(\wigglyComplex_3) & = (1, 19, 68, 68, 19, 1) \\
h(\wigglyComplex_4) & = (1, 48, 420, 1147, 1147, 420, 48, 1) \\
h(\wigglyComplex_5) & = (1, 109, 1960, 11254, 25868, 25868, 11254, 1960, 109, 1)
\end{align*}

\end{slide}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{slide}{Wiggly pseudotriangulations}

$c$ cell in a wiggly dissection with boundary $\partial_c$ \\
\emph{degree} $\delta_c =  1/2 \, \# \text{arcs on } \partial_c + 2 \, \# \text{connected components of } \partial_c - 1$ \\
\emph{pseudotriangle} $=$ cell of degree $3$ \qquad \emph{pseudoquadragle} $=$ cell of degree $4$

\vspace{.4cm}
\centerline{\includegraphics[scale=3]{wigglyPseudodissections}}

\vspace{.1cm}
\theo{PROP}{The inclusion maximal wiggly pseudodissections are the pseudotriangulations, and contain $2n-1$ internal arcs and $n$ cells.}

\vspace{.3cm}
%$wp_n = \#$wiggly pseudotriangulations
\[
\begin{array}{c|ccccccccc}
n & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \dots \\
\hline
wp_n & 2 & 14 & 176 & 3232 & 78384 & 2366248 & 85534176 & 3602770400 & \dots
\end{array}
\]

\end{slide}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{slide}{Wiggly flip graph}

\theo{PROP}{Any wiggly pseudoquadrangle has exactly two wiggly diagonals, and they either cross or are non pointed.}

\vspace{.1cm}
\centerline{\includegraphics[scale=2]{diagonalsPseudoquadrangles}}

\end{slide}

%%%

\begin{slide}{Wiggly flip graph}

\theo{PROP}{Any wiggly pseudoquadrangle has exactly two wiggly diagonals, and they either cross or are non pointed.}

\vspace{.1cm}
\centerline{\includegraphics[scale=2]{diagonalsPseudoquadrangles}}

\vspace{-.4cm}
\emph{wiggly flip graph} $\wigglyFlipGraph_n =$ \begin{tabular}[t]{|ll} a vertex for each wiggly pseudotriangulation \\ an edge between~$T$ and~$T'$ if~$T \ssm \{\alpha\} = T' \ssm \{\alpha'\}$ \end{tabular}

\vspace{.3cm}
\centerline{\includegraphics[scale=3]{wigglyFlip}}

\end{slide}

%%%

\begin{slide}{Wiggly flip graph}

\vspace{.5cm}
\theo{PROP}{The wiggly flip graph $\wigglyFlipGraph_n$ is $(2n-1)$-regular and connected.}

\vspace{.5cm}
\centerline{\includegraphics[scale=1.7]{wigglyComplex} \hspace{-2cm} \includegraphics[scale=1.7]{wigglyComplexDual}}

\end{slide}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\partie{Wiggly permutations \& wiggly lattice}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{slide}{Wiggly permutations}

\vspace{1cm}
\hspace{13.5cm}
\includegraphics[scale=2]{wigglyLattice}

\vspace{-13.5cm}
\emph{wiggly permutation} = permutation of $2n$ avoiding
\begin{compactitem}
\item $(2j-1) \cdots i \cdots (2j)$ for~$j \in [n]$ and~$i < 2j-1$,
\item $(2j) \cdots k \cdots (2j-1)$ for~$j \in [n]$ and~$k > 2j$.
\end{compactitem}

\vspace{1cm}
\begin{minipage}{13.2cm}
\theo{PROP}{
The wiggly permutations induce a sublattice~$\wigglyLattice_n$ of the weak order on~$\f{S}_{2n}$.

}

\theo{PROP}{
The cover graph of the lattice~$\wigglyLattice_n$ is $(2n-1)$-regular and connected.
}
\end{minipage}

\vspace{5cm}
\[
\begin{array}{c|ccccccccc}
n & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \dots \\
\hline
wp_n & 2 & 14 & 176 & 3232 & 78384 & 2366248 & 85534176 & 3602770400 & \dots
\end{array}
\]

\end{slide}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{slide}{Bijection}

\vspace{1cm}
\centerline{\includegraphics[scale=1.42]{wigglyPseudotriangulations}}
\vspace{-.3cm}
\centerline{$
1234 \;\;\; 1243 \;\;\; 1342 \;\;\; 1423 \;\;\; 1432 \;\;\; 2134 \;\;\; 2143 \;\;\; 3412 \;\;\; 3421 \;\;\; 4123 \;\;\; 4132 \;\;\; 4213 \;\;\; 4312 \;\;\; 4321
$}

\vspace{1cm}
\theo{PROP}{The wiggly pseudotriangulations and wiggly permutations are in bijection.}

\vspace{.5cm}
\centerline{\includegraphics[scale=3]{bijection}}

\end{slide}

%%%

\begin{slide}{Bijection}

\theo{PROP}{This bijection induces a directed graph isomorphism between
\begin{compactitem}
\item the wiggly increasing flip graph on wiggly pseudotriangulations,
\item the Hasse diagram of the wiggly lattice on wiggly permutations.
\end{compactitem}
}

\vspace{.5cm}
\centerline{\includegraphics[scale=1.8]{wigglyFlipGraph}\quad\raisebox{.5cm}{\includegraphics[scale=1.8]{wigglyLattice}}}

\end{slide}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\partie{Wiggly fan}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{slide}{g- and c-vectors}

\emph{$\b{g}$-vector} of $\alpha =$ proj. on $\sum_{i = 1}^{2n} x_i = 0$ of \raisebox{-1.7cm}{\includegraphics[scale=4]{gvector}}

\vspace{-.5cm}
\emph{$\b{c}$-vector} of~$\alpha \in T = $ you don't want to know...

\vspace{.5cm}
\theo{PROP}{For any wiggly pseudotriangulation~$T$, the $\b{g}$-vectors~$\set{\b{g}(\alpha)}{\alpha \in T^\circ}$ and the $\b{c}$-vectors~$\set{\b{c}(\alpha, T)}{\alpha \in T^\circ}$ form dual bases.}

\vspace{1cm}
\centerline{\raisebox{-4cm}{\includegraphics[scale=4]{pseudotriangulationMatrices}} \quad \input{figures/gcmatrices}}

\end{slide}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{slide}{wiggly fan}

\emph{$\b{g}$-vector} of $\alpha =$ proj. on $\sum_{i = 1}^{2n} x_i = 0$ of \raisebox{-1.7cm}{\includegraphics[scale=4]{gvector}}

\vspace{.1cm}
\theo{THM}{The cones~$\setangle{\b{g}(\alpha)}{\alpha \in D}$ for all wiggly dissections~$D$ form a complete simplicial fan~$\wigglyFan_n$ (in $\sum_{i = 1}^{2n} x_i = 0$).
}

\vspace{2cm}
Main observation:

\centerline{\includegraphics[scale=3.2]{incompatible2}}
\centerline{
${\red \b{g}(\alpha)} + {\orange \b{g}(\alpha')} = \big( {\green \b{g}(\beta)} + {\blue \b{g}(\beta')} \big) / 2$
\qquad
${\red \b{g}(\alpha)} + {\orange \b{g}(\alpha')} = {\green \b{g}(\beta)} + {\blue \b{g}(\beta')}$
}

\end{slide}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\partie{Wigglyhedron}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{slide}{Wigglyhedron}

\emph{incompatibility degree}~$\delta(\alpha, \alpha') =$ 
\begin{compactitem}
\item $0$ if~$\alpha$ and~$\alpha'$ are pointed and non-crossing,
\item $1$ is~$\alpha$ and~$\alpha'$ are not pointed,
\item the number of crossings of~$\alpha$ and~$\alpha'$ if they are crossing.
\end{compactitem}

\vspace{.5cm}
$\kappa(\alpha) =$ \emph{incompatibility number} of~$\alpha = \sum_{\alpha'} \delta(\alpha, \alpha')$.

\vspace{2cm}
Main observation:

\centerline{\includegraphics[scale=3.2]{incompatible2}}
\centerline{
${\red \kappa(\alpha)} + {\orange \kappa(\alpha')} > \big( {\green \kappa(\beta)} + {\blue \kappa(\beta')} \big) / 2$
\qquad
${\red \kappa(\alpha)} + {\orange \kappa(\alpha')} > {\green \kappa(\beta)} + {\blue \kappa(\beta')}$
}

\vspace{1.5cm}
Hence, $\kappa$ satisfies all wall-crossing inequalities of the wiggly fan...

\end{slide}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{slide}{Wigglyhedron}

\theo{THM}{
The wiggly fan~$\wigglyFan_n$ is the normal fan of a simplicial $(2n-1)$-dimensional polytope, called the \emph{wigglyhedron}~$\wigglyhedron_n$, and defined equivalently~as
\begin{itemize}
\item intersection of the halfspaces~$\set{\b{x} \in \R^{2n}}{\dotprod{\b{g}(\alpha)}{\b{x}} \!\le\! \kappa(\alpha)}$ for all wiggly arcs~$\alpha$,
\item convex hull of~$\b{p}(T) \eqdef \sum\limits_{\alpha \in T} \kappa(\alpha) \, \b{c}(\alpha, T)$ for all wiggly pseudotriangulations~$T$.
\end{itemize}
\vspace{-.5cm}
}

\vspace{1cm}
\centerline{\input{figures/wigglyhedron}}

\end{slide}

%%%

\begin{slide}{Wigglyhedron}

\theo{THM}{The wigglyhedron~$\wigglyhedron_n$ is a simple $(2n-1)$-dimensional polytope such that
\begin{compactitem}
\item the wiggly complex~$\wigglyComplex_n$ is the boundary complex of the polar of~$\wigglyhedron_n$,
\item the Hasse diagram of the wiggly lattice is a linear orientation of the graph of~$\wigglyhedron_n$.
\end{compactitem}
}

\vspace{1cm}
\centerline{\includegraphics[scale=1.7]{wigglyFlipGraph} \quad \input{figures/wigglyhedron}}

\end{slide}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\partie{Cambrian considerations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{slide}{Cambrian considerations}

\theo{THM}{
For $\delta = \delta_1 \dots \delta_n \in \{+,-\}^n$, there are lattice isomorphisms between 
\begin{itemize}
\item \emph{$\delta$-triangulations} $=$ triangulation of the $\delta$-gon, whose vertex at abscissa $i$ has ordinate \\[-1cm]
\[
\text{positive if $\delta_j = {+}$
\qquad and \qquad
negative if $\delta_j = {-}$}
\]
\vspace{-1.5cm}
\item \emph{$\delta$-permutations} $=$ permutation of~$[n]$ avoiding for~$i < j < k$ \\[-1cm]
\[
\cdots j \cdots ki \cdots \text{ if~$\delta_j = +$}
\quad\text{and}\quad
\cdots ik \cdots j \cdots \text{ if~$\delta_j = -$}
\]
\vspace{-1.5cm}
\item \emph{$\delta$-wiggly pseudotriangulations} $=$ wiggly pseudotriangulation containing the arcs \\[-1cm]
\[
(0, j, {[1,j[}, \varnothing) \text{ for~$\delta_j = {+}$}
\qquad\text{and}\qquad
(0, j, \varnothing, {[1,j[}) \text{ for~$\delta_j = {-}$}
\]
\vspace{-1.5cm}
\item \emph{$\delta$-wiggly permutations} $=$ wiggly permutation~$\sigma$ of~$[2n]$ such that \\[-1cm]
\[
\delta_j = {+} \quad \Longrightarrow \quad \sigma^{-1}(i) \le \sigma^{-1}(2j-1)
\quad\text{and}\quad
\delta_j = {-} \quad \Longrightarrow \quad \sigma^{-1}(2j) \le \sigma^{-1}(i)\quad\;\;
\]
\end{itemize}
\vspace{-.5cm}
}

\vspace{.5cm}
\centerline{\includegraphics[scale=1.6]{wigglyCambrian1}}

\end{slide}

%%%

\begin{slide}{Cambrian considerations}

\vspace{.5cm}
\centerline{\includegraphics[scale=1.6]{wigglyCambrian2}}

\end{slide}

%%%

\begin{slide}{Cambrian considerations}

\vspace{.5cm}
\centerline{\includegraphics[scale=1.6]{wigglyCambrian3}}

\end{slide}

%%%

\begin{slide}{Cambrian considerations}

\theo{THM}{
For $\delta = \delta_1 \dots \delta_n \in \{+,-\}^n$, there are lattice isomorphisms between 
\begin{itemize}
\item \emph{$\delta$-triangulations} $=$ triangulation of the $\delta$-gon, whose vertex at abscissa $i$ has ordinate \\[-1cm]
\[
\text{positive if $\delta_j = {+}$
\qquad and \qquad
negative if $\delta_j = {-}$}
\]
\vspace{-1.5cm}
\item \emph{$\delta$-permutations} $=$ permutation of~$[n]$ avoiding for~$i < j < k$ \\[-1cm]
\[
\cdots j \cdots ki \cdots \text{ if~$\delta_j = +$}
\quad\text{and}\quad
\cdots ik \cdots j \cdots \text{ if~$\delta_j = -$}
\]
\vspace{-1.5cm}
\item \emph{$\delta$-wiggly pseudotriangulations} $=$ wiggly pseudotriangulation containing the arcs \\[-1cm]
\[
(0, j, {[1,j[}, \varnothing) \text{ for~$\delta_j = {+}$}
\qquad\text{and}\qquad
(0, j, \varnothing, {[1,j[}) \text{ for~$\delta_j = {-}$}
\]
\vspace{-1.5cm}
\item \emph{$\delta$-wiggly permutations} $=$ wiggly permutation~$\sigma$ of~$[2n]$ such that \\[-1cm]
\[
\delta_j = {+} \quad \Longrightarrow \quad \sigma^{-1}(i) \le \sigma^{-1}(2j-1)
\quad\text{and}\quad
\delta_j = {-} \quad \Longrightarrow \quad \sigma^{-1}(2j) \le \sigma^{-1}(i)\quad\;\;
\]
\end{itemize}
\vspace{-.5cm}
}

\vspace{.5cm}
\theo{PROP}{
The $\delta$-associahedron~$\Asso_\delta$ is normally equivalent to the face of the wigglyhedron~$\wigglyhedron_n$ corresponding to the wiggly pseudodissection formed by the $\delta$-wiggly arcs.

}

\end{slide}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\partie{Some open problems}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{slide}{Open problem 1: graph properties of wigglyhedron}

\centerline{\includegraphics[scale=1.7]{wigglyFlipGraph}\quad\includegraphics[scale=1.7]{wigglyLattice}}

\theo{QU 1a}{Is the wiggly flip graph Hamiltonian?

\hfill \color{gray}{[NB: wiggly permutations do not form a zigzag language, right?]}
}

\theo{QU 1b}{What is the diameter of the wiggly flip graph?

%\hfill \color{gray}{[NB: $2n-1 \le \delta_n$]}
}

\end{slide}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{slide}{Open problem 2: wiggly pseudotriangulations of point sets}

$\b{P}$ arbitrary point set in the plane

\emph{wiggly complex}~$\wigglyComplex_{\b{P}}$ = simplicial complex of non-crossing and pointed wiggly edges

\vspace{-.5cm}\hspace{14.5cm}
\rotatebox{270}{\includegraphics[scale=1.1]{wigglyComplexSquare}}

\vspace*{-15cm}
\begin{minipage}{13cm}
\theo{Q2a}{Is~$\wigglyComplex_{\b{P}}$ the boundary complex of a simplicial polytope?

\color{gray}{
[NB: aligned points $\Rightarrow$ wigglyhedron \\
general position $\Rightarrow$ Rote--Santos--Streinu]
}

\vspace{.1cm}
}

\vspace{1cm}
\theo{Q2b}{Is the graph of~$\wigglyComplex_{\b{P}}$ Hamiltonian?}

\end{minipage}

\end{slide}

\end{document}